### 2022-02-14 acks/acks.tex
initial

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
\setheader{Preface}

Without a doubt, the acknowledgments are the most widely and most
eagerly read part of any thesis.

\begin{flushright}
{\makeatletter\itshape
    Petar \\
    Amsterdam, April 2022
\makeatother}
\end{flushright}