### 2022-02-15 ehoh/ehoh.tex
toc

\chapter{Extending a Brainiac Prover to Lambda-Free Higher-Order Logic}
\setheader{Extending a Brainiac Prover to Lambda-Free Higher-Order Logic}
\label{ch:ehoh}