### 2022-02-23 ehoh/ehoh.tex
index

\chapter{Extending a Brainiac Prover to Lambda-Free Higher-Order Logic}
\setheader{Extending a Brainiac Prover to Lambda-Free Higher-Order Logic}
\label{ch:ehoh}

\DeclareMathAlphabet{\mathcalx}{OT1}{pzc}{m}{it}
\renewcommand{\confrep}[2]{#2}
\includeversion{rep}
\excludeversion{conf}

\blfootnote{In this work I desgined, implemented and evaluated all changes to term representation, algorithms and indexing data structures.}

\authors{Joint work with\\
Jasmin~Blanchette, Simon~Cruanes and Stephan~Schulz}

\begin{abstract}
Decades of work have gone into developing efficient proof calculi, data
structures, algorithms, and heuristics for first-order automatic theorem
proving. Higher-order provers lag behind in terms of efficiency. Instead of
developing a new higher-order prover from the ground up, we propose to start
with the state-of-the-art superposition prover E and gradually enrich it with
higher-order features. We explain how to extend the prover's data structures,
algorithms, and heuristics to $\lambda$-free higher-order logic, a formalism
that supports partial application and applied variables. Our extension
outperforms the traditional encoding and appears promising as a stepping stone
toward full higher-order logic.    
\end{abstract}

\newpage

\section{Introduction}
\label{sec:ehoh:introduction}

\begin{sloppypar}
Superposition provers such as E~\cite{scv-19-e23}, SPASS \cite{wdfksw-09-spass},
and Vampire \cite{lkav-13-vampire} are among the most successful first-order
reasoning systems. They serve as backends in various frameworks, including
software verifiers (e.g., Why3 \cite{fp-13-why3}),
%
% \pagebreak[2]%
automatic higher-order theorem provers (e.g., \hbox{Leo-III} \cite{sb-21-leo3},
Satallax \cite{cb-12-satallax}), and \begin{rep}one-click \end{rep}``hammers'' in proof assistants
(e.g., HOLyHammer in HOL Light \cite{ku-15-holyhammer}, Sledgehammer in
Isabelle \cite{pb-12-sh}).

Dec\-ades of research have gone
into refining calculi, devising efficient data structures and algorithms,
and developing heuristics to guide proof search\begin{rep}
  \cite{ss-17-arcade}\end{rep}.
This work has mostly focused on first-order logic with equality.
\end{sloppypar}

Research on higher-order automatic provers has resulted
in systems such as LEO \cite{cbmk-98-leo}, \textsc{Leo}-II
\cite{bspt-15-leo2}, and Leo-III \cite{sb-21-leo3},
based on resolution and paramodulation, and Satallax \cite{cb-12-satallax},
based on \confrep{tableaux}{tableaux and SAT solving}. They
feature a ``cooperative'' architecture, pioneered by LEO: They
are full-fledged higher-order provers that regularly invoke an
external first-order prover \confrep{}{with a low time limit as a terminal
procedure, }in an attempt to finish the proof quickly using only
first-order reasoning. However, the first-order backend will succeed
only
if all the necessary higher-order reasoning has been performed,
meaning that much of the first-order reasoning is carried out by the
slower higher-order prover. As a result, this architecture leads to
suboptimal performance on largely first-order problems,
such as those that often arise
in interactive verification \cite{ns-13-leo2sh}. For example,
at the 2017 installment of the CADE ATP System Competition (CASC)
\cite{gs-17-casc}, Leo-III, which uses E as a backend, proved
652 out of 2000 first-order problems in the Sledgehammer division, compared
with 1185 for E on its own and 1433 for Vampire.

\looseness=-1
To obtain better performance, we propose to start with a competitive
first-order prover and extend it to full higher-order logic one
feature at a time.  Our goal is a \emph{graceful} extension, so that
the system behaves as before on first-order problems, performs mostly
like a first-order prover on typical, mildly higher-order problems,
and scales up to arbitrary higher-order problems, in keeping with the
zero-overhead principle: \emph{What you don't use, you
  don't pay for.}

As a stepping stone toward full higher-order logic, we initially restrict our
focus to a higher-order logic without \hbox{$\lambda$-expressions}
(Sect.~\ref{sec:ehoh:logic}). Compared with first-order logic, its distinguishing
features are partial application and applied variables. It
is rich enough to express the recursive equations of higher-order
combinators, such as $\cst{map}$ on lists:
%
\begin{align*}
\cst{map} \> f \> \cst{nil} & \eq \cst{nil}
&
\cst{map} \> f \> (\cst{cons} \; x \; \mathit{xs}) & \eq
\cst{cons} \; (f \> x) \; (\cst{map} \> f \> \mathit{xs})
\end{align*}

Our vehicle is E\begin{rep} \cite{ss-02-brainiac,scv-19-e23}\end{rep},
a prover developed primarily by Schulz. It is written in C and offers
good performance, with more emphasis on ``brainiac''
heuristics than on raw speed. E regularly scores among the top
systems at CASC and is usually the strongest open-source prover in
the relevant divisions. It also serves as a backend for
competitive higher-order provers. We refer to our extended version of
E as Ehoh. It corresponds to a prerelease version of E~2.5 configured with the
option \verb|--enable-ho|.%
\footnote{\url{https://github.com/eprover/eprover/commit/80946ac}}
%We plan to include the new higher-order inferences in the E 2.5 release.

The main challenges we faced concerned the
representation of types and terms
(Sect.~\ref{sec:ehoh:types-and-terms}), the unification
\confrep{algorithm}{and matching algorithms}
(Sect.~\ref{sec:ehoh:unif-match}), and the indexing data structures
(Sect.~\ref{sec:ehoh:indexing}). We also adapted the
inference rules (Sect.~\ref{sec:ehoh:inferences})\confrep{ and}{,} the
heuristics (Sect.~\ref{sec:ehoh:heuristics})\confrep{}{, and the
  preprocessor (Sect.~\ref{sec:ehoh:preprocessing})}.

\looseness=-1
A central aspect of our work is a set of techniques we call
\emph{prefix optimization}. Taking a traditional look at higher-order terms, they contain twice as many proper
subterms as first-order terms; for example,
$\cst{f}\;(\cst{g}\;\cst{a})\;\cst{b}$ contains not only the ``argument'' subterms
$\cst{g}\;\cst{a}$, $\cst{a}$, $\cst{b}$ but also the ``prefix'' subterms
$\cst{f}$, $\cst{f}\;(\cst{g}\;\cst{a})$, $\cst{g}$.
\begin{rep}Many operations, including superposition and rewriting, require
traversing all subterms of a term.\end{rep}
Using the optimization, the prover traverses subterms recursively in a
first-order fashion, considering all the prefixes of a given subterm
together. %, at little additional cost.
%
Our experiments (Sect.~\ref{sec:ehoh:evaluation}) show that Ehoh is
almost as fast as E on first-order problems and can also prove
higher-order problems that do not require synthesizing
$\lambda$-terms. As next steps, we plan to add support for
$\lambda$-terms and higher-order unification.

\section{Logic}
\label{sec:ehoh:logic}

\looseness=-1
Our logic is a variant of the intensional $\lambda$-free Boolean-free higher-order
logic (\lfhol{}) described by Bentkamp et al.\ %, Blanchette, Cruanes, and Waldmann
\cite[Sect.~2]{bbcw-21-lfho}, which could also be called ``applicative
first-order logic.'' In the spirit of FOOL \cite{kotelnikov-16-fool}, we
%% kotelnikov-15-fool
extend the syntax of this logic by erasing the distinction between terms and
formulas, and its semantics by interpreting the Boolean type $o$ as a domain
of cardinality~2. Functional extensionality can be obtained by adding
suitable axioms \cite[Sect.~3.1]{bbcw-21-lfho}.

\looseness=-1

This logic differs from the higher-order logic described in
Sect.~\ref{sec:pre:hol} three ways. First, $\lambda$-abstraction is disallowed.
Second, logical connectives are not part of the set of symbols. Instead, there
is a special inductive case in the definition of terms that defines formulas.
Third, subterms are defined in a more traditional way.

\looseness=-1
For completeness, we provide the definition of terms. Terms, ranged over by $s,t,u,v$, are either
\emph{variables} $x, y, z, \dots$, (\emph{function}) \emph{symbols} $\cst{a},
\cst{b}, \cst{c}, \cst{d},\allowbreak \cst{f},\allowbreak \cst{g}, \ldots{}$
(often called ``constants'' in the higher-order literature), binary applications
$s \; t$, or Boolean terms $\itrue$, $\ifalse$, $\inot s$, $s \mathbin{\iand}
t$, $s \mathbin{\ior} t$, $s \mathbin{\iimplies} t$, $s \mathbin{\iequiv} t$,
$\iforall x.\> s$, $\iexists x.\> s$, $s \mathrel{\ieq} t$. E and Ehoh clausify
the input as a preprocessing step, producing a clause set in which the only
proper Boolean subterms are variables, $\itrue$, and $\ifalse$. Note that we use
lowercase letters for free variables as bound variables do not appear in
clauses. A term's \emph{arity} is the number of extra arguments it can take. If
$\iota$ is a base type, $\cst{f}$ has type $\iota \to \iota \to \iota$, and
$\cst{a}$ has type $\iota$, then $\cst{f}$ is binary, $\cst{f}\;\cst{a}$ is
unary, and $\cst{f}\;\cst{a}\;\cst{a}$ is nullary. Subterms are defined in the
traditional higher-order way; for example, $s \; t$ has all subterms of $s$ and
$t$ as subterms, in addition to $s \; t$ itself; as a consequence
$\cst{f} \; \cst{a}$ is a subterm of $\cst{f} \; \cst{a} \;\cst{a}$. With $\Var(x)$
we denote the set of free variables of $x$ where $x$ ranges over terms, clauses,
set of clauses or any other objects that contain terms.

In this chapter, substitutions $\sigma$ are partial
functions of finite domain from variables to terms, written $\{ x_1 \mapsto s_1,
\ldots, x_m \mapsto s_m \}$, where each $s_i$ has the same type as $x_i$. The
substitution $\sigma[x \mapsto s]$ maps $x$ to $s$ and otherwise coincides with
$\sigma$. Applying $\sigma$ to a variable beyond $\sigma$'s domain is the
identity. We deviated from the view of substitutions in Section \ref{sec:pre:hol}
as it made proofs in Sections \ref{sec:ehoh:unif-match} and \ref{sec:ehoh:indexing}
easier. It is easy to check that both views are equivalent. We also consider unification 
constraints $s \unif t$ as \emph{ordered} pairs.
%Applying a substitution to a term~$t$ applies it homomorphically to $t$'s
%variables.

A well-known technique to support \lfhol{} %in first-order reasoning systems
is to use the \emph{applicative encoding}:
Every $n$-ary symbol is mapped to a nullary symbol, and
application is represented by a distinguished binary symbol $\cst{@}.$ Thus,
the \lfhol{} term $\cst{f} \; (x\; \cst{a}) \; \cst{b}$ is
encoded as the first-order term $\cst{@}(\cst{@}(\cst{f}, \cst{@}(x,
\cst{a})), \cst{b}).$ However, this representation is not graceful, since it
also introduces $\cst{@}$'s for terms within \lfhol's first-order fragment. By
doubling the size and depth of terms, the encoding clutters data
structures and slows down term traversals.
%By changing the root symbol of terms,
%it impacts proof search heuristics in subtle ways.
In our empirical evaluation, we find that the applicative encoding can
decrease the success rate by up to \NumberOK{15\%} (Sect.~\ref{sec:ehoh:evaluation}).
For these and further reasons, it
is not ideal (Sect.~\ref{sec:ehoh:discussion-and-related-work}).

\section{Types and Terms}
\label{sec:ehoh:types-and-terms}

\looseness=-1
The term representation is a central concern when building a theorem
prover. Delicate changes to E's representation were needed to support
partial application and especially applied variables. In contrast, the
introduction of a higher-order type system had a less dramatic impact on the
prover's code.
%In this section we
%describe changes we made to generalize E's term representation and how
%we completely replaced FOL many-sorted type system with simple types
%needed for \lfhol{}.

\ourpara{Types} For most of its history, E supported only untyped first-order logic. Cruanes
implemented support for atomic types for E 2.0
\cite[p.\,117]{sc-15-simon-phd}. Symbols are declared
with a type signature:
$\cst{f} : (\tau_1, \ldots, \tau_{n}) \rightarrow \tau.$
Atomic types are represented by integers, leading to efficient type comparisons.

\looseness=-1
In \lfhol{}, a type signature is simply a type~$\tau$, in which the
type constructor $\to$ can be nested---e.g., $(\iota \to \iota)\allowbreak
\to \iota.$
%
A natural way to represent such types is to mimic their recursive
structure using a tagged union. However, this leads to memory fragmentation;
a simple operation such as querying the type of a function's $i$th
argument would require dereferencing $i$ pointers. We prefer a
flattened representation, in
which a type $\tau_1 \rightarrow \cdots \rightarrow \tau_n \rightarrow \iota$
is represented by a single node labeled with ${\rightarrow}$ and pointing to
the array $(\tau_1,\dots,\tau_n,\iota).$

Ehoh stores all types in a shared bank and
implements perfect sharing, ensuring that types that are structurally the same
are represented by the same object in memory. Type equality can then be
implemented as a pointer comparison.

\ourpara{Terms}

In E, terms are stored as perfectly shared directed acyclic graphs\begin{rep}
\cite{ls-01-shared}\end{rep}.
%
%\begin{rep}
%-- or kill completely
%Even though we could implement applicative encoding as a preprocessor outside of E
%and feature full support for \lfhol{} in that way, we decided to generalize existing
%data structures and algorithms and relax existing invariants that are valid
%only for FOL terms. These generalizations and relaxations are largely dependent
%on the way E deals with term representation.
%\end{rep}
%
Each node, or \emph{cell}, contains 11~fields, including
\verb|f_code|, an integer that identifies the term's head symbol (if ${\ge}\;0$)
or variable (if ${<}\;0$); \verb|arity|, an integer corresponding to the number
of arguments passed to the head; \verb|args|, an array of size
\verb|arity| consisting of pointers to arguments; and \verb|binding|,
which may store a substitution for a variable\begin{rep} (if
\verb|f_code|$\;{<}\;0$)\end{rep}, used for unification and matching.

In first-order logic, the arity of a variable is always 0, and the arity of a
symbol is given by its type signature.
In higher-order logic, variables may have function type and be applied, and
symbols can be applied to fewer arguments than specified by their type
signatures. A natural representation of \lfhol{} terms as tagged unions
would distinguish between variables~$x$, symbols~$\cst{f}$, and binary
applications $s \; t.$ However, this scheme suffers from memory
fragmentation and linear-time access, as with the representation of types,
affecting performance on purely or mostly first-order problems. Instead, we
propose a flattened representation, as a generalization of E's existing data
structures: Allow arguments to variables, for symbols let \verb|arity| be
the number of \begin{conf}\emph{actual} arguments\end{conf}\begin{rep}actual
arguments, %as opposed to the declared arity,
and rename the field \verb|num_args|\end{rep}. This representation, often called
``spine notation,'' % \cite{cervesato-pfenning-2003},
is isomorphic to the standard definition of higher-order terms with binary
application. It is employed in various higher-order reasoning systems,
including Leo-III \cite{sb-21-leo3} and
Zipperposition \cite{bbtvw-21-sup-lam}.

%\begin{rep} This approach parallels our representation of types.\end{rep}

A side effect of the flattened representation is that prefix subterms are not
shared. For example, the terms $\cst{f}\;\cst{a}$ and $\cst{f}\;\cst{a}\;\cst{b}$
correspond to the flattened cells $\cst{f}(\cst{a})$ and
$\cst{f}(\cst{a}, \cst{b}).$ The argument subterm $\cst{a}$ is shared, but not
the prefix $\cst{f}\;\cst{a}.$ Similarly,
$x$ and $x\;\cst{b}$ are represented by two distinct cells, $x()$ and
$x(\cst{b})$, and there is no connection between the two occurrences of~$x$.
%
In particular,
despite perfect sharing, their \verb|binding| fields are unconnected, leading
to inconsistencies.

A potential solution would be to systematically traverse a
clause and set the \verb|binding| fields of all cells of the form $x(\overline{s})$
whenever a variable~$x$ is bound, but this would be inefficient and inelegant.
Instead, we implemented a hybrid approach: Variables are applied by an
explicit application operator
\appvar{}, to ensure that they are always perfectly shared. Thus, $x\;\cst{b}\;\cst{c}$
is represented by the cell $\appvar(x, \cst{b}, \cst{c})$, where $x$ is a shared
subcell. This is graceful, since variables never occur applied in first-order
terms.
%
The main drawback is that some normalization is necessary
after substitution: Whenever a variable is instantiated by a symbol-headed term,
the $\appvar$ symbol must be eliminated.
Applying the substitution $\{x \mapsto \cst{f} \; \cst{a}\}$ to the cell
$\appvar(x, \cst{b}, \cst{c})$ must produce $\cst{f}(\cst{a}, \cst{b}, \cst{c})$ and
not $\appvar(\cst{f}(\cst{a}), \cst{b}, \cst{c})$, for consistency with other
occurrences of $\cst{f} \; \cst{a} \; \cst{b} \; \cst{c}.$

There is one more complication related to the \verb|binding| field. In E, it
is easy and useful to traverse a term as if a substitution has been applied,
by following all set \verb|binding| fields. In Ehoh, this is not enough,
because cells must also be normalized. To avoid repeatedly creating the
same normalized cells, we introduced a \verb|binding_cache| field
that connects a $\appvar(x, \overline{s})$ cell with its substitution.
%
However, this cache can easily become stale when $x$'s \verb|binding| pointer is
updated. To detect this situation, we store $x$'s \verb|binding|
value in the $\appvar(x, \overline{s})$ cell's \verb|binding|
field (which is otherwise unused).
To find out whether the cache is valid, it suffices to check that the
\verb|binding| fields of $x$ and $\appvar(x, \overline{s})$ are equal.

\ourpara{Term Orders} Superposition provers rely on term orders to prune the search
space. The order must be a simplification order that
%can be extended to a simplification order that
% That's true but pointless. It's equivalent to saying that you need a
% simplification order and the prover may underapproximate it. -JB
is total on variable-free
terms. E implements both the Knuth--Bendix order (KBO) and the lexicographic path order
(LPO). KBO is widely regarded as the more
robust option for superposition.
%
In earlier work, Blanchette and colleagues have shown that only KBO can be
generalized gracefully while preserving the necessary properties for
superposition \cite{bbww-17-kbo,bww-17-rpo}. For this
reason, we focus on KBO.

E implements L\"ochner's linear-time algorithm for KBO
\cite{bl-06-kbo}, which relies on the \relax{tupling} method to store
intermediate results. % , avoiding repeated computations.
It is straightforward to generalize the algorithm to compute the
graceful \lfhol{} version of KBO \cite{bbww-17-kbo}.
The main difference is that when comparing two terms $\cst{f} \; \tuple{s}{m}$
and $\cst{f} \; \tuple{t}{n}$, because of partial application we may now
have $m \not= n$; this required changing the implementation to
perform a length-lexicographic comparison of the tuples $\tuple{s}{m}$ and
$\tuple{t}{n}.$

\ourpara{Input and Output Syntax} 
E implements the TPTP \cite{gs-17-tptp} formats FOF and TF0, corresponding to
untyped and mono\-morphic first-order logic, for both input and output. In
Ehoh, we added support for the \lfhol{} fragment of TPTP TH0, which provides
mono\-morphic higher-order logic. Thanks to the use of a standard format,
Ehoh's proofs can immediately be parsed by Sledgehammer
\cite{pb-12-sh}, which reconstructs them using a variety of
techniques.
There is ongoing work on increasing the level of detail of E's proofs, to
facilitate proof interchange and independent proof checking
\cite{rs-17-checkable};
this will also benefit Ehoh.

\section{Unification and Matching}
\label{sec:ehoh:unif-match}

Syntactic unification of \lfhol{} terms has a first-order flavor.
It is decidable, and most general unifiers (MGUs) are unique up to variable
renaming. For example, the unification constraint $\cst{f} \; (x \;
\cst{a}) \unif x \; (\cst{f} \; \cst{a})$, used to illustrate an infinite set of 
independent unifiers in full higher-order logic (Sect.~\ref{sec:pre:unif}), has the MGU 
$\{x \mapsto \cst{f}\}$ in \lfhol{}.
% in \lfhol{} the unification constraint $\cst{f} \; (y \;
% \cst{a}) \unif y \; (\cst{f} \; \cst{a})$ has the MGU $\{y \mapsto \cst{f}\}$,
% while in full higher-order logic infinitely many independent solutions exist as
% described in Sect.~\ref{sec:pre:unif}. 
Matching is a special case of unification
where only the variables on the left-hand side can be instantiated.

An easy but inefficient way to implement unification and matching for \lfhol{}
is to apply the applicative encoding (Sect.~\ref{sec:ehoh:logic}), perform
first-order unification or matching, and decode the resulting
substitution. To avoid the overhead, we generalize the first-order unification
and matching procedures to operate directly on \lfhol{} terms.

\subsection{Unification}
We present our unification procedure as a \begin{rep}nondeterministic \end{rep}transition system
that generalizes Baader and Nipkow \cite{bn-98-tr-and-all-that}.
A unification problem consists of a finite set~$\mathit{S}$ of unification constraints $s_i
\unif t_i$, where $s_i$ and $t_i$ are of the same type.
A problem is in \emph{solved form} if it has the form $\{ x_1 \unif t_1,\allowbreak
\ldots, x_n \unif t_n \}$, where the $x_i$'s are distinct and do not occur in
the $t_j$'s. The corresponding unifier is
$\{ x_1 \mapsto t_1,\allowbreak \ldots, x_n \mapsto t_n \}.$
%
The transition rules attempt to bring the input constraints into solved form.
\begin{rep}They can be applied in any order and eventually reach a normal form, which is
either an idempotent MGU expressed in solved form or the special
value $\bot$, denoting unsatisfiability of the constraints.\end{rep}

The first group of rules\confrep{ }{---the \emph{positive} rules---}consists
of operations that focus on a single constraint and replace it with a new
(possibly empty) set of constraints:

\noindent
\begin{tabular}{ll}
  \textsf{Delete} & $\{ t \unif t \} \uplus \mathit{S} \Unifarrow \mathit{S}$ \\[\jot]
  \textsf{Decompose} & $\{ \cst{f} \; \tuple{s}{m} \unif \cst{f} \; \tuple{t}{m} \} \uplus \mathit{S} \Unifarrow {} 
                        \mathit{S} \cup \{ s_1 \unif t_1, \ldots, s_m \unif t_m \}$ \\[\jot]
  \textsf{DecomposeX} & $\{ x \; \tuple{s}{{m}} \unif u \; \tuple{t}{{m}} \} \uplus \mathit{S} \Unifarrow {} 
                         \mathit{S} \cup \{ x \unif u{,}\; s_1 \unif t_1, \ldots, s_{m} \unif t_{m} \}$  \\
                      & if $x$ and $u$ have the same type and $m > 0$ \\[\jot]
  \textsf{Orient}     & $\{ \cst{f} \; \overline{s} \unif x \; \overline{t} \} \allowbreak \uplus \mathit{S}
                         \Unifarrow \mathit{S} \cup \{ x \; \overline{t} \unif \cst{f} \; \overline{s} \}$ \\[\jot]
  \textsf{OrientXY}   &  $\{ x \; \tuple{s}{m} \unif y \; \tuple{t}{n} \} \allowbreak \uplus \mathit{S}
                         \Unifarrow \mathit{S} \cup \{ y \; \tuple{t}{n} \unif x \; \tuple{s}{m} \}$  \\
                      &   if $m > n$ \\[\jot]
  \textsf{Eliminate}  & $\{ x \unif t \} \uplus \mathit{S} \Unifarrow \{ x \unif t \} \cup \{ x \mapsto t \}(\mathit{S})$ \\
                      &   if $x \in \Var(\mathit{S}) \setminus \Var(t)$
\end{tabular}

The \textsf{Delete}, \textsf{Decompose}, and \textsf{Eliminate} rules are
essentially as for first-order terms. The \textsf{Orient} rule is generalized
to allow applied variables and complemented by a new \textsf{Orient\-XY} rule.
\textsf{DecomposeX}, also a new rule, can be seen as a variant of
\textsf{Decompose} that analyzes applied variables; the term $u$ may be an
application.

The rules of the second group\confrep{ }{---the \emph{negative}
rules---}detect unsolvable constraints:
%
% \begin{description}[labelwidth=\widthof{\rm\textsf{OccursCheck}}]
% \item[\rm\textsf{Clash}]
%   $\{ \cst{f} \; \overline{s} \unif \cst{g} \; \overline{t} \} \uplus \mathit{S} \Unifarrow \bot$%
%   \quad if $\cst{f} \not= \cst{g}$

% \smallskip
% \item[\rm\textsf{ClashTypeX}]
%   \begin{tabular}[t]{@{}l@{}}
%     $\{ x \; \tuple{s}{m} \unif u \; \tuple{t}{m} \} \allowbreak \uplus \mathit{S} \Unifarrow \bot$ \\
%     if $x$ and $u$ have different types
%   \end{tabular}

% \smallskip
% \item[\rm\textsf{ClashLenXF}]
%   $\{ x \; \tuple{s}{m} \unif \cst{f} \; \tuple{t}{n} \} \allowbreak \uplus \mathit{S} \Unifarrow \bot$%
%   \quad if $m > n$

% \smallskip
% \item[\rm\textsf{OccursCheck}]
%   $\{ x \unif t \} \uplus \mathit{S} \Unifarrow \bot$%
%   \quad if $x \in \Var(t)$ and $x \not= t$
% \end{description}
%

\noindent
\begin{tabular}{ll}
  \textsf{Clash} &   $\{ \cst{f} \; \overline{s} \unif \cst{g} \; \overline{t} \} \uplus \mathit{S} \Unifarrow \bot$; if $\cst{f} \not= \cst{g}$ \\[\jot]
  \textsf{ClashTypeX} & $\{ x \; \tuple{s}{m} \unif u \; \tuple{t}{m} \} \allowbreak \uplus \mathit{S} \Unifarrow \bot$ ; if $x$ and $u$ have different types \\[\jot]
  \textsf{ClashLenXF} & $\{ x \; \tuple{s}{m} \unif \cst{f} \; \tuple{t}{n} \} \allowbreak \uplus \mathit{S} \Unifarrow \bot$; if $m > n$ \\[\jot]
  \textsf{OccursCheck}     & $\{ x \unif t \} \uplus \mathit{S} \Unifarrow \bot$;  if $x \in \Var(t)$ and $x \not= t$
\end{tabular}

\begin{rep}\textsf{Clash} and \textsf{OccursCheck} are essentially as in
Baader and Nipkow. \textsf{ClashTypeX} and \textsf{ClashLenXF}
are variants of \textsf{Clash} for applied variables.\end{rep}

\newcommand\unifARROW[1]{\rlap{\ensuremath{\Unifarrow_{#1}\;}}\phantom{\Unifarrow_\textsf{DecomposeX}\;}}

The derivation below demonstrates the computation of MGUs for the
unification problem
$\{x \> (z \> \cst{b} \> \cst{c}) \unif \cst{g} \> \cst{a} \> (y \> \cst{c})\}$:
%
\begin{align*}
& \{ \relax{x \> (z \> \cst{b} \> \cst{c})  \unif  \cst{g} \>  \cst{a} \>  (y \> \cst{c})}\}
\\[-1\jot]
\unifARROW{\textsf{DecomposeX}}
& \{ \relax{x \unif \cst{g} \> \cst{a}}{,}\; z \> \cst{b} \> \cst{c} \unif y \> \cst{c} \}
\\[-1\jot]
\unifARROW{\textsf{OrientXY}}
& \{ x \unif \cst{g} \> \cst{a}{,}\; \relax{y \> \cst{c} \unif z \> \cst{b} \> \cst{c}} \}
\\[-1\jot]
\unifARROW{\textsf{DecomposeX}}
& \{ x \unif \cst{g} \> \cst{a}{,}\; \relax{y  \unif z \> \cst{b}}{,}\;  \cst{c} \unif  \cst{c} \}
\\[-1\jot]
\unifARROW{\textsf{Delete}}
& \{ x \unif \cst{g} \> \cst{a}{,}\; y  \unif z \> \cst{b} \}
\end{align*}

E stores open constraints in a double-ended
queue. Constraints are processed from the front. New constraints are
added at the front if they involve complex
terms that can be dealt with swiftly by \textsf{Decompose} or
\textsf{Clash}, or to the back if one side is a variable. \begin{rep}This
delays instantiation of variables %(with a possible increase in term size)
and allows E to detect structural clashes early. \end{rep}%

During proof search, E repeatedly needs to test a term~$s$ for unifiability
not only with some other term $t$ but also with $t$'s subterms.
Prefix optimization speeds up this test: The subterms of $t$ are
traversed in a first-order fashion; for each such subterm $\zeta \;
\tuple{t}{n}$, at most one prefix $\zeta \;
\tuple{t}{k}$, with $k \le n$, is possibly unifiable with $s$, by virtue
of their having the same arity. \begin{rep}For first-order problems, we can only have $k =
n$, since all functions are fully applied. \end{rep}Using this technique, Ehoh
is virtually as efficient as E on first-order terms.

\looseness=-1
  The transition system introduced above always terminates with a correct
  answer. Our proofs follow the lines of Baader and Nipkow.
  The metavariable $\mathcalx{R}$ is used to range over constraint sets $\mathit{S}$
  and the special value $\bot.$
  The set of all unifiers of $\mathit{S}$ is denoted by $\Unifier(\mathit{S}).$
  Note that $\Unifier(\mathit{S} \cup \mathit{S}') = \Unifier(\mathit{S}) \cap \Unifier(\mathit{S}').$
  We let $\Unifier(\bot) = \emptyset.$
  The notation $\mathit{S} \Unifarrow^{!} \mathit{S}'$ indicates that $\mathit{S}
  \Unifarrow^{*} \mathit{S}'$ and $\mathit{S}'$ is a normal form (i.e., there exists no
  $\mathit{S}''$ such that $\mathit{S}' \Unifarrow \mathit{S}''$).
  A variable $x$ is \emph{solved} in $\mathit{S}$
  if it occurs exactly once in $\mathit{S}$, in a constraint of the form $x \unif t.$

  \begin{lemma}\label{lemma:ehoh:unif-preserve}
    If $\mathit{S} \Unifarrow \mathcalx{R}$, then $\Unifier(\mathit{S}) = \Unifier(\mathcalx{R}).$
  \end{lemma}
  \begin{proof}
    The rules \textsf{Delete}, \textsf{Decompose}, \textsf{Orient}, and
    \textsf{Eliminate} are proved as in Baader and Nipkow.
    \textsf{OrientXY} trivially preserves unifiers.
    For \textsf{DecomposeX}, the core of the argument is as follows:
    %
    \begin{align*}
    & \sigma \in \Unifier(\{ x \; \tuple{s}{{m}} \unif u \; \tuple{t}{{m}} \}) \\[-.5\jot]
      \text{iff}\quad
    & \sigma(x \; \tuple{s}{{m}}) = \sigma(u \; \tuple{t}{{m}}) \\[-.5\jot]
      \text{iff}\quad
    & \sigma(x) \; \sigma(s_1) \, \ldots \, \sigma(s_m)
        = \sigma(u) \; \sigma(t_1) \, \ldots \, \sigma(t_m) \\[-.5\jot]
      \text{iff}\quad
    & \sigma(x) = \sigma(u){,}~ \sigma(s_1) = \sigma(t_1){,}\; \ldots{,}~ \text{and}~ \sigma(s_m) = \sigma(t_m) \\[-.5\jot]
      \text{iff}\quad
    & \sigma \in \Unifier(\{ x \unif u{,}\; s_1 \unif t_1, \ldots, s_{m} \unif t_{m}\})
    \end{align*}

    The proof of the problem's unsolvability if rule \textsf{Clash} or
    \textsf{OccursCheck} is applicable carries over from Baader and Nipkow.
    For \textsf{ClashTypeX}, the justification is that $\sigma(x \;
    \tuple{s}{m}) = \sigma(u \; \tuple{t}{m})$ is possible only if
    $\sigma(x) = \sigma(u)$, which requires $x$ and $u$ to have the same
    type. Similarly, for \textsf{ClashLenXF}, if
    $\sigma(x \; \tuple{s}{m}) = \sigma(\cst{f} \; \tuple{t}{n})$ with
    $m > n$, we must have $\sigma(x \; \tuple{s}{{m-n}}) =
    \sigma(x) \; \sigma(s_1) \, \dots \, \sigma(s_{m-n}) = \cst{f}$,
    which is impossible.
  \end{proof}
\newpage
  \begin{lemma}\label{lemma:ehoh:unif-normal-imp-solved}
    If $\mathit{S}$ is a normal form, then $\mathit{S}$ is in solved form.
  \end{lemma}
  \begin{proof}
    Consider an arbitrary unification constraint $s \unif t \in \mathit{S}.$ We show
    that in all but one cases, a rule is applicable, contradicting the
    hypothesis that $\mathit{S}$ is a normal form. In the remaining case, $s$ is a
    solved variable in $\mathit{S}$.

    \medskip
    \noindent
    \textsc{Case} $s = x$:
    \begin{itemize}
    \item \textsc{Subcase} $t = x$:
      \textsf{Delete} is applicable.

%      \smallskip
    \item \textsc{Subcase} $t \neq x$ and $x \in \Var(t)$:
      \textsf{OccursCheck} is applicable.

%      \smallskip
    \item \textsc{Subcase} $t \neq x$, $x \not \in \Var(t)$, and $x \in
      \Var(\mathit{S} \setminus \{ s \unif t \})$:
      \textsf{Eliminate} is applicable.

%      \smallskip
    \item \textsc{Subcase} $t \neq x$, $x \not \in \Var(t)$,
          and $x \notin \Var(\mathit{S} \setminus \{ s \unif t \})$:
      The variable $x$ is solved in $\mathit{S}$.
    \end{itemize}

    \noindent
    \textsc{Case} $s = x \; \tuple{s}{m}$ for $m > 0$:
    \begin{itemize}
    \item \textsc{Subcase} $t = \eta \; \tuple{t}{n}$ for $n \ge m$:
      \textsf{DecomposeX} or \textsf{Clash\-TypeX} is applicable, depending on
      whether $x$ and $\eta \; \tuple{t}{{n-m}}$ have the same type.

%      \smallskip
    \item \textsc{Subcase} $t = y \; \tuple{t}{n}$ for $n < m$:
      \textsf{OrientXY} is applicable.

%      \smallskip
    \item \textsc{Subcase} $t = \cst{f} \; \tuple{t}{n}$ for $n < m$:
      \textsf{ClashLenXF} is applicable.
    \end{itemize}

    \noindent
    \textsc{Case} $s = \cst{f} \; \tuple{s}{m}$:
    \begin{itemize}
    \item \textsc{Subcase} $t = x \; \tuple{t}{n}$:
      \textsf{Orient} is applicable.

%      \smallskip
    \item \textsc{Subcase} $t = \cst{f} \; \tuple{t}{n}$:
      Due to well-typedness, $m = n$. \textsf{Decompose} is
      applicable.

%      \smallskip
    \item \textsc{Subcase} $t = \cst{g} \; \tuple{t}{n}$:
      \textsf{Clash} is applicable.
    \end{itemize}

    \noindent
    Since each constraint is of the form $x \unif t$ where $x$ is solved in
    $\mathit{S}$, the problem $\mathit{S}$ is in solved form.
  \end{proof}

  \begin{lemma}\label{lemma:ehoh:unif-idempotent-mgu}
    If the constraint set $\mathit{S}$ is in solved form, then the associated
    substitution is an idempotent MGU of $\mathit{S}.$
  \end{lemma}
  \begin{proof}
    This lemma corresponds to Lemma~4.6.3 of Baader and Nipkow. Their proof
    carries over to \lfhol{}.
  \end{proof}

  \begin{theorem}[Partial Correctness]\label{thm:ehoh:unif-partial-correctness}\,%
    If $\mathit{S} \Unifarrow^{!} \bot$, then $\mathit{S}$ has no solutions.
    If $\mathit{S} \Unifarrow^{!} \mathit{S}'\!$, then $\mathit{S}'$ is in solved form and the
    associated substitution is an idempotent MGU of $\mathit{S}$.
    \end{theorem}

    \begin{proof}
    The first part follows from Lemma~\ref{lemma:ehoh:unif-preserve}.
    %applied transitively
    The second part follows from Lemma \ref{lemma:ehoh:unif-preserve}
    %applied transitively
    and Lemmas
    \ref{lemma:ehoh:unif-normal-imp-solved}~and~\ref{lemma:ehoh:unif-idempotent-mgu}.
    \end{proof}

      \newpage
      \begin{theorem}[Termination]\label{thm:ehoh:unif-termination}\,%
      The relation $\Unifarrow$ is well founded.
      \end{theorem}

      \begin{proof}
      We define an auxiliary notion of weight:
      $\mathcalx{W}(\zeta \; \tuple{s}{m}) =
        \allowbreak m + 1 + \sum_{i=1}^{m} \mathcalx{W}(s_i)$.
      Well-foundedness is proved by exhibiting a measure function from
      constraint sets to quadruples of natural numbers $(n_1, n_2, n_3, n_4)$,
      where
      $n_1$ is the number of unsolved variables in $\mathit{S}$;
      $n_2$ is the sum of all term weights,
      $\sum_{s \unif t \in \mathit{S}} \mathcalx{W}(s) + \mathcalx{W}(t)$;
      $n_3$ is the number of right-hand sides with variable heads,
      $|\{ s \unif x \; \overline{t} \in \mathit{S} \} |$; and
      $n_4$ is the number of arguments to left-hand side variable heads,
        $\sum_{x \> \tuple{s}{m} \unif t  \in \mathit{S}} m$.

      The following table shows that the application of each positive rule
      lexicographically decreases the quadruple:

        \begin{center}
          \begin{tabular}{l@{\qquad}l@{\quad}l@{\quad}l@{\quad}l@{\quad}}
                              & $n_1$  & $n_2$ & $n_3$ & $n_4$ \\[\jot]
          \textsf{Delete}     & $\geq$ & $>$   &       &       \\
          \textsf{Decompose}  & $\geq$ & $>$   &       &       \\
          \textsf{DecomposeX} & $\geq$ & $>$   &       &       \\
          \textsf{Orient}     & $\geq$ & $=$   & $>$   &       \\
          \textsf{OrientXY}   & $\geq$ & $=$   & $=$   & $>$   \\
          \textsf{Eliminate}  & $>$    &       &       &
          \end{tabular}
        \end{center}
  %
  %    \noindent
  %    \textsf{Eliminate}, by solving one variable, decreases $n_1.$
  %    \textsf{Delete} removes a constraint from~$\mathit{S}$, decreasing the
  %    weight.
  %    \textsf{Decompose} reduces the weight by at least~2.
  %    \textsf{Prefix} reduces the set weight by at least $m>0$, thanks to the
  %    presence of $m$ in the definition of $\mathcalx{W}$.
      The negative rules, which produce the special value $\bot$, cannot
      contribute to an infinite $\Unifarrow$ chain.
      %For example, the quadruples corresponding to the derivation
      %starting from $\{x \> (z \> \cst{b} \> \cst{c}) \unif \cst{g} \> \cst{a} \> (y \> \cst{c})\}$
      %given above are $(3,14,0,1) > (2,12,1,2) > (2, 12, 1, 1) > (1, 10, 1, 0) > (1, 8, 1, 0)$.
  %    where $>$ denotes the lexicographic extension of $>$ on natural numbers.
      \end{proof}

  A unification algorithm for \lfhol{} can be derived from the above transition
  system, by committing to a strategy for applying the rules.
  %
  This algorithm closely follows the Ehoh implementation, abstracting away from
  complications such as prefix optimization. We assume a flattened
  representation of terms; as in Ehoh, each variable stores the term it is bound
  to in its \textit{binding} field (Sect.~\ref{sec:ehoh:types-and-terms}).
  We also rely on a \textsc{ApplySubst} function, which applies the
  binding to the top-level variable. The algorithm assumes that the terms
  to be unified have the same type. The pseudocode is as follows:

  \begin{quotex}
    \MyFunction{SwapNeeded}{Term $\mathit{s}$, Term $\mathit{t}$}
    \q \MyReturn $\begin{aligned}[t]
      & t.\mathit{head}.\mathit{isVar}() \\[-\jot]
      & {\mathrel\land}\; (\lnot\, s.\mathit{head}.\mathit{isVar}()
      \mathrel\lor s.\mathit{num\_args} > t.\mathit{num\_args})
      \end{aligned}$

    \vskip 1.5\jot

    \MyFunction{Deref}{Term $\mathit{s}$}
    \q \MyWhile{$s.\mathit{head}.\mathit{isVar}() \mathrel\land s.\mathit{head}.\mathit{binding} \neq \mathit{Null}$}
    \qq $s \gets \textsc{ApplySubst}(s{,}\; s.\mathit{head}.\mathit{binding})$ \\
    \q \MyReturn $s$

    \vskip 1.5\jot

    \MyFunction{GobblePrefix}{Term $x$, Term $t$}
    \q $\mathit{res} \gets \mathit{Null}$ \\
    \q \MyIf{$x.\mathit{type}.\mathit{args}$ is suffix of
          $\mathit{t}.\mathit{head}.\mathit{type}.\mathit{args}$}
    \qq $\mathit{pref\_len} \gets \mathit{t}.\mathit{head}.\mathit{type}.\mathit{arity} - x.\mathit{type}.\mathit{arity}$ \\
    \qq \MyIf{$\mathit{pref\_len} \leq t.\mathit{num\_args}$}
    \qqq $\mathit{res} \gets $
                \textsc{Term}$(t.\mathit{head}$, $t.\mathit{args}[1 \,.\,.\, \mathit{pref\_len}])$ \\
    \q \MyReturn $\mathit{res}$

    \vskip 1.5\jot

    \MyFunction{Unify}{Term $\mathit{s}$, Term $\mathit{t}$}
    \q $\mathit{constraints} \gets \textsc{DoubleEndedQueue}()$ \\
    \q $\mathit{constraints}.\mathit{prepend(s)}$ \\
    \q $\mathit{constraints}.\mathit{prepend(t)}$
    \\[\jot]
    \q \MyWhile{$\lnot\, \mathit{constraints}.\mathit{isEmpty}()$}
    \qq $t \gets \textsc{Deref}(\mathit{constraints}.\mathit{dequeue}())$ \\
    \qq $s \gets \textsc{Deref}(\mathit{constraints}.\mathit{dequeue}())$
    \\[\jot]
    \qq \MyIf{$s \neq t$}
    \qqq \MyIf{\textsc{SwapNeeded}$(s,t)$}
    \qqqq $(t, s) \gets (s, t)$
    \\[\jot]
    \qqq \MyIf{$s.\mathit{head}.\mathit{isVar}()$}
    \qqqq $x \gets s.\mathit{head}$ \\
    \qqqq $\mathit{prefix} \gets \textsc{GobblePrefix}(x,t)$ \\
    \qqqq \MyIf{$\mathit{prefix} \neq \mathit{Null}$}
    \qqqqq $\mathit{start\_idx} \gets \mathit{prefix}.\mathit{num\_args} + 1$ \\
    \qqqqq \MyIf{$x$ occurs in $\mathit{prefix}$}
    \qqqqqq \MyReturn $\textit{False}$ \\
    \qqqqq \MyElse
    \qqqqqq $x.\mathit{binding} \gets \mathit{prefix}$ \\
    \qqqq \MyElse
    \qqqqq \MyReturn $\textit{False}$ \\
    \qqq \MyElsIf{$s.\mathit{head} = t.\mathit{head}$}
    \qqqq  $\mathit{start\_idx} \gets 1$ \\
    \qqq \MyElse
    \qqqq \MyReturn $\textit{False}$
    \\[\jot]
    \qqq \MyForTo{$i \gets \mathit{start\_idx}$}{$t.\mathit{num\_args}$}
    \qqqq $\mathit{s\_arg} \gets s.\mathit{args}[i-\mathit{start\_idx}+1]$ \\
    \qqqq $\mathit{t\_arg} \gets t.\mathit{args}[i]$ \\
    \qqqq \MyIf{$(\mathit{s\_arg}.\mathit{head}.\mathit{isVar}() \lor \mathit{t\_arg}.\mathit{head}.\mathit{isVar}())$}
    \qqqqq $\mathit{constraints}.\mathit{append}(t\_\mathit{arg})$ \\
    \qqqqq $\mathit{constraints}.\mathit{append}(s\_\mathit{arg})$ \\
    \qqqq \MyElse
    \qqqqq $\mathit{constraints}.\mathit{prepend}(s\_\mathit{arg})$ \\
    \qqqqq $\mathit{constraints}.\mathit{prepend}(t\_\mathit{arg})$
    \\[\jot]
    \q \MyReturn $\textit{True}$
    \end{quotex}

\subsection{Matching}

Given $s$ and $t$, the matching problem consists of finding a
substitution $\sigma$ such that $\sigma(s) = t.$
We then write that ``$t$ is an instance of $s$'' or
``$s$~generalizes~$t$.'' %and call $\sigma$ a \emph{generalizing substitution}.
We are interested in most general generalizations (MGGs).
Matching can be reduced to unification by treating variables in $t$ as
nullary symbols \cite{bn-98-tr-and-all-that}, but E implements it
separately.

Matching can be specified abstractly as a transition system on matching constraints
$s_i \MATCH t_i$ consisting of the
unification rules \textsf{Decompose},
\textsf{DecomposeX},
\textsf{Clash},
\textsf{ClashTypeX},
\textsf{ClashLenXF}
(with $\MATCH$ instead of $\unif$)
and augmented with

\noindent
\begin{tabular}{ll}
  \textsf{Double} &   $\{ x \MATCH t{,}\; x \MATCH t' \} \uplus \mathit{S} \Matcharrow \bot$; if $t \not= t'$  \\[\jot]
  \textsf{ClashLenXY} & $\{ x \; \tuple{s}{m} \MATCH y \; \tuple{t}{n} \} \allowbreak \uplus \mathit{S} \Matcharrow \bot$; if $x \not= y$ and $m > n$ \\[\jot]
  \textsf{ClashFX} & $\{ \cst{f} \; \overline{s} \MATCH x \; \overline{t} \} \uplus \mathit{S} \Matcharrow \bot$
\end{tabular}

The matching relation is sound, complete, and well founded. Interestingly, a
\textsf{Delete} rule would be unsound for matching. Consider the problem
$\{ x \MATCH x, x \MATCH \cst{g} \; x \}$. Applying \textsf{Delete} to the
first constraint would yield the solution $\{ x \MATCH \cst{g} \; x \}$, even
though the original problem is clearly unsolvable.

\section{Indexing Data Structures}
\label{sec:ehoh:indexing}

Superposition provers like E work by saturation. Their main loop
heuristically selects a clause and searches for potential inference
partners among a possibly large set of other clauses. Mechanisms such
as simplification and subsumption also require locating terms in a
large clause set. For example, when E derives a new equation
$s \eq t$, if $s$ is larger than $t$ according to the term order, it
will rewrite all instances $\sigma(s)$ of $s$ to $\sigma(t)$ in
existing clauses.

To avoid iterating over all terms (including subterms) in large clause
sets, superposition provers store the potential inference partners in
indexing data structures. A term index stores a set of terms $\mathit{S}.$
Given a \emph{query term}
%
$t$, a query returns all terms $s \in \mathit{S}$ that satisfy a given
\emph{retrieval condition}:\ $\sigma(s) = \sigma(t)$ ($s$~and $t$ are
unifiable), $\sigma(s) = t$ ($s$ generalizes~$t$), or $s = \sigma(t)$
($s$ is an instance of $t$), for some substitution~$\sigma.$
\emph{Perfect} indices return exactly the subset of terms satisfying
the retrieval condition. In contrast, \emph{imperfect} indices return
a superset of eligible terms, and the retrieval condition needs to be
checked for each candidate.

E relies on two term indexing data structures, perfect discrimination
trees \cite{mcc-92-pdts} and fingerprint indices
\cite{ss-12-fp-indexing}, that needed to be generalized to
\lfhol. It also uses feature vector indices
\cite{ss-2013-feature-vector} to speed up subsumption and
related techniques, but these require no changes to work
with \lfhol{}.

\subsection{Discrimination Trees}

\looseness=-1
Discrimination trees \cite{mcc-92-pdts} are tries in which every
node is labeled with a symbol or a variable. A path from the root to
a leaf corresponds to a ``serialized term''---a term expressed without
parentheses and commas. Consider the following discrimination trees $D_1$ and
$D_2$:
%
\begin{align*}
 & \vcenter{\hbox{\begin{tikzpicture}[level distance=2.25em, sibling distance=5em,
%  edge from parent path={(\tikzparentnode.south) -- ++(0,-0.5em)
%   -| (\tikzchildnode.north)},
  every node/.style = {shape=circle, minimum size=1.75em, draw, align=center}]
  \node {\phantom{x}}
    child { node {$\cst{f}$}
      child { node %[fill=verylightgray]
        {$\cst{a}$} }
    }
    child { node % [fill=verylightgray]
      {$\cst{g}$}
      child { node {$\cst{a}$}
        child { node % [fill=verylightgray]
          {$\cst{a}$} }
      }
      child[sibling distance=2.5em] { node {$\cst{b}$}
        child { node %[fill=verylightgray]
          {$\cst{a}$} }
        child { node %[fill=verylightgray]
          {$\cst{b}$} }
      }
    }
    ;
\end{tikzpicture}
}}
&&
  \vcenter{\hbox{\begin{tikzpicture}[level distance=2.25em, sibling distance=5em,
%  edge from parent path={(\tikzparentnode.south) -- ++(0,-0.5em)
%   -| (\tikzchildnode.north)},
  every node/.style = {shape=circle,
    draw, align=center}]
  \node {\phantom{x}}
    child { node {$\cst{f}$}
      child { node %[fill=verylightgray]
        {$x$} }
    }
    child { node % [fill=verylightgray]
      {$\cst{g}$}
      child { node {$\cst{a}$}
        child { node % [fill=verylightgray]
          {$\cst{a}$} }
      }
      child[sibling distance=2.5em] { node {$y$}
        child { node %[fill=verylightgray]
          {$\cst{a}$} }
        child { node %[fill=verylightgray]
          {$x$} }
      }
    }
    child[sibling distance=3.75em] { node {$x$} }
    ;
\end{tikzpicture}
}}
\end{align*}
%
Assuming $\cst{a}, \cst{b}, x, y : \iota$, $\cst{f} : \iota \to \iota$,
and $\cst{g} : \iota^2 \to \iota$,
$D_1$ represents the term set
$\{
\cst{f}(\cst{a})
{,}\allowbreak\;
\cst{g}(\cst{a}, \cst{a})
{,}\allowbreak\;
\cst{g}(\cst{b}, \cst{a})
{,}\allowbreak\;
\cst{g}(\cst{b}, \cst{b})
\}$,
and $D_2$ represents the term set
$\{
\cst{f}(x)
{,}\allowbreak\;
\cst{g}(\cst{a}, \cst{a})
{,}\allowbreak\;
\cst{g}(y, \cst{a})
{,}\allowbreak\;
\cst{g}(y, x)
{,}\;
x
\}$.
%
E uses perfect discrimination trees for finding generalizations of
query terms. Thus, if the query term is
$\cst{g}(\cst{a}, \cst{a})$, it would follow the path $\cst{g}.\cst{a}.\cst{a}$
in~$D_1$ and return $\{\cst{g}(\cst{a}, \cst{a})\}.$
For~$D_2$, it would also explore paths labeled with
variables, binding them as it proceeds, and return $\{
\cst{g}(\cst{a}, \cst{a}){,}\allowbreak\; \cst{g}(y, \cst{a}){,}\allowbreak\; \cst{g}(y, x){,}\; x \}.$

It is crucial for this data structure that distinct terms always give rise to
distinct serialized terms. Conveniently, this property also holds for \lfhol{}
terms. Suppose that two distinct \lfhol{} terms yield the same serialization.
Clearly, they must disagree on parentheses; one will have the subterm $s\; t\;
u$ where the other has $s\; (t\; u).$ However, these two subterms cannot both
be well typed.

\looseness=-1
When generalizing the data structure to \lfhol, we face a complication
due to partial application. First-order terms can only be stored in leaf nodes,
but in Ehoh we must also be able to represent partially applied terms,
such as $\cst{f}$, $\cst{g}$, or $\cst{g}\;\cst{a}$ (assuming, as above, that
$\cst{f}$ is unary and $\cst{g}$ is binary). Conceptually, this can be solved
by storing a Boolean on each node indicating whether it is an accepting state.
In the implementation, the change is more subtle, because several parts of E's
code implicitly assume that only leaf nodes are accepting.

\newcommand\TermStack{\ensuremath{T}}
\newcommand\BTStack{\ensuremath{P}}

The main difficulty specific to \lfhol{} concerns applied variables. To
enumerate all generalizing terms, E needs to backtrack from child to
parent nodes. This is achieved using two stacks that store subterms of
the query term:
\TermStack{} %(\texttt{term\_\allowbreak stack})
stores the terms that
must be matched in turn against the
current subtree, and
\BTStack{} %(\texttt{term\_\allowbreak proc})
stores, for each node from
the root to the current subtree, the corresponding processed term.

Let $[a_1, \dots, a_n]$ denote an $n$-item stack with $a_1$ on top. Given a
query term $t$, the matching procedure starts at the root with $\sigma =
\emptyset$, $\TermStack{} = [t]$, and $\BTStack = []$.
%
The procedure advances by repeatedly moving to a suitable child node:
%
\begin{enumerate}
\item[A.] If the node is labeled with a symbol $\cst{f}$ and the top item
  $t$ of \TermStack{} is \confrep{}{of the form }$\cst{f}(\tuple{t}{n})$,
  replace $t$ by $n$~new items $t_1,\dots,t_n$, and push $t$ onto \BTStack.

\smallskip
\item[B.] If the node is labeled with a variable $x$, there are two subcases.
  If $x$ is already bound, check that $\sigma(x) = t$;
  otherwise, extend $\sigma$ so that $\sigma(x) = t.$ Next, pop the term $t$
  from \TermStack{} and push it onto \BTStack.
\end{enumerate}

%
The goal is to reach an accepting node. If the query term and all the terms
stored in the tree are first-order, \TermStack{} will then be empty, and the
entire query term will have been matched.
%
Backtracking works in reverse: Pop a term $t$ from \BTStack; if the current
node is labeled with an $n$-ary symbol, discard \TermStack{}'s topmost
$n$~items; push $t$ onto \TermStack. Undo any variable bindings.

As an example, looking up $\cst{g}(\cst{b}, \cst{a})$ in the tree $D_1$
would result in the following succession of stack states, starting from the
root $\varepsilon$ along the path $\cst{g}.\cst{b}.\cst{a}$:
%
\[\begin{array}{@{}l@{\kern1.5em}l@{\kern1em}l@{\kern1em}l@{\kern1em}l@{}}
& \varepsilon & \cst{g} & \cst{g}.\cst{b} & \cst{g}.\cst{b}.\cst{a} \\[.5\jot]
\sigma{:} & \emptyset & \emptyset & \emptyset & \emptyset \\
\TermStack{:} & [\cst{g}(\cst{b}, \cst{a})] & [\cst{b}, \cst{a}] & [\cst{a}] & [\,] \\
\BTStack{:} & [\,] & [\cst{g}(\cst{b}{,}\; \cst{a})] & [\cst{b}{,}\; \cst{g}(\cst{b}{,}\; \cst{a})] & [\cst{a}{,}\; \cst{b}{,}\; \cst{g}(\cst{b}, \cst{a})]
\end{array}\]
%
Backtracking amounts to moving leftward:
To get back from $\cst{g}$ to the root, we pop $\cst{g}(\cst{b}, \cst{a})$
from \BTStack, we discard two items from \TermStack{}, and we push
$\cst{g}(\cst{b}, \cst{a})$ onto \TermStack{}.

To adapt the procedure to \lfhol{}, the key idea is that an applied variable
is not very different from an applied symbol. A node labeled with an $n$-ary
head~$\zeta$ matches a prefix~$t'$ of the $k$-ary term $t$
popped from \TermStack{} and leaves $n - k$ arguments $\overline{u}$ to be
pushed back, with $t = t' \; \overline{u}.$ If $\zeta$ is a variable, it must
be bound to the prefix~$t'$ assuming $\zeta$ and $t'$ are of same type.
%
Backtracking works analogously: Given the arity $n$ of the node label $\zeta$
and the arity $k$ of the term $t$ popped from \BTStack{}, we discard the
topmost $n - k$ items $\overline{u}$ from \BTStack.

\looseness=-1
To illustrate the procedure, we consider the tree $D_2$ but change $y$'s type
to $\iota \to \iota.$ This tree stores
$\{
\cst{f}\; x
{,}\allowbreak\;
\cst{g}\; \cst{a}\; \cst{a}
{,}\allowbreak\;
\cst{g}\; (y\; \cst{a})
{,}\allowbreak\;
\cst{g}\; (y\; x)
{,}\allowbreak\;
x
\}.$
Let $\cst{g}\; (\cst{g}\;\cst{a}\;\cst{b})$ be the query term.
We have the following sequence of substitutions $\sigma$ and stacks
$\TermStack, \BTStack$:
%
\[\begin{array}{@{}l@{\kern1em}l@{\kern1em}l@{\kern1em}l@{}}
\varepsilon & \cst{g} & \cst{g}.y & \cst{g}.y.x \\[.5\jot]
\emptyset & \emptyset & \{y \mapsto \cst{g}\> \cst{a}\} & \{y \mapsto \cst{g}\> \cst{a}{,}\; x \mapsto \cst{b}\} \\[0pt]
[\cst{g}\> (\cst{g}\> \cst{a}\> \cst{b})] & [\cst{g}\> \cst{a}\> \cst{b}] & [\cst{b}] & [\,] \\[0pt]
[\,] & [\cst{g}\> (\cst{g}\> \cst{a}\> \cst{b})] & [\cst{g}\> \cst{a}\> \cst{b}{,}\;\cst{g}\> (\cst{g}\> \cst{a}\> \cst{b})] & [\cst{b}{,}\;\cst{g}\> \cst{a}\> \cst{b}{,}\;\cst{g}\> (\cst{g}\> \cst{a}\> \cst{b})]
\end{array}\]

\confrep{}{When backtracking from $\cst{g}.y$ to $\cst{g}$, by comparing
$y$'s arity of $n = 1$ with $\cst{g}\; \cst{a}\; \cst{b}$'s arity of $k = 0$,
we determine that one~item must be discarded from \TermStack{}. }%
Finally, to avoid traversing twice as many subterms as in the first-order
case, we can optimize prefixes: Given a
query term $\zeta\;\tuple{t}{n}$, we can also match prefixes
$\zeta\;\tuple{t}{k}$, where $k < n$, by allowing \TermStack{} to be nonempty when we reach an accepting node.
%If accepting nodes correspond to legal prefixes, the stack will then contain
%exactly the remaining arguments: $[t_{k+1},\dots,t_n].$

\newcommand\quadruple{\mathcalx{Q}}
Similarly to unification and matching, we present finding generalizations
in a perfect discrimination tree as a transition system. States
are quadruples $\quadruple = (\overline{t}, \overline{b}, \mathit{D}, \sigma)$, where
$\overline{t}$ is a list of terms, $\overline{b}$ is a list of tuples storing
backtracking information, $\mathit{D}$ is a discrimination
(sub)tree, and $\sigma$ is a substitution.

%With $+$ we denote list append operation, and with
%$::$ we denote operation of prepending single element to list.

Let $D$ be a perfect discrimination tree. $\Term(D)$ denotes the set of terms
stored in $D$. The function $D|_\zeta$ returns the child of $D$ labeled with
$\zeta$, if it exists. Child nodes are themselves perfect discrimination
(sub)trees. Given any node $D$, if the node is accepting, then the value stored
on that node is defined as $\mathit{val}(D) = (s, d)$, where $s$ is the
accepted term and $d$ is some arbitrary data;
%(typically, an equation whose one side is~$s$)
otherwise, $\mathit{val}(D)$ is undefined.

%We label the tree nodes using BFS, starting with 0 as label for
%root node.

%With $\mathcalx{R}$ we denote $(\overline{t}, \overline{b}, D, \sigma)$ or
%$(\mathit{v}, \sigma)$ or $\bot$.

Starting from an initial state $([t], [\,], D, \emptyset)$, where $t$ is the query
term and $D$ is an entire discrimination tree, the following transitions are
possible:
% \begin{description}[labelwidth=\widthof{\rm\textsf{AdvanceX}}]
%   \item[\rm\textsf{AdvanceF}]
%     $ ( \cst{f} \> \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma )
%         \Pdtarrow
%         (\tuple{s}{m} \cdot \overline{t}{,}\;
%         (\cst{f} \> \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_\cst{f}{,}\; \sigma)
%       $\quad
%       if $D{|}_\cst{f}$ is defined

%   \smallskip
%   \item[\rm\textsf{AdvanceX}]
%     $ ( s \; \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma )
%         \Pdtarrow {}$ \\
%     $(\tuple{s}{m} \cdot \overline{t}{,}\; (s \; \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_x{,}\; \sigma[x \mapsto s])
%       $\\
%       if $D{|}_x$ is defined, $x$ and $s$ have the same type, and
%       $\sigma(x)$ is either undefined or equal to $s$

%   \smallskip
%   \item[\rm\textsf{Backtrack}]
%     $ ( \tuple{s}{m} \cdot \overline{t}{,}\; (s, D_0, \sigma_0) \cdot \overline{b}{,}\; D{,}\; \sigma)
%         \Pdtarrow %\\\hphantom{\rm\textsf{AdvanceX}}%
%       (s \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D_0{,}\; \sigma_0)$ \\
%       if $D_0{|}_\zeta = D$ and $m = \textit{arity}(\zeta) - \textit{arity}(s)$

%   \smallskip
%   \item[\rm\textsf{Success}]
%     $ ( [\,]{,}\; \overline{b}{,}\; D{,}\; \sigma )
%         \Pdtarrow (\mathit{val}(D){,}\; \sigma)$ \\
%     if~$\mathit{val}(D)$ is defined
% \end{description}

\noindent
\begin{tabular}{ll}
  \textsf{AdvanceF} & $ ( \cst{f} \> \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma )
           \Pdtarrow
           (\tuple{s}{m} \cdot \overline{t}{,}\;
           (\cst{f} \> \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_\cst{f}{,}\; \sigma)
         $ \\
                   & if $D{|}_\cst{f}$ is defined \\[\jot]
  \textsf{AdvanceX} & $( s \; \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma ) 
                      \Pdtarrow {} 
                      (\tuple{s}{m} \cdot \overline{t}{,}\; (s \; \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_x{,}\; \sigma[x \mapsto s])$ \\
                    & if $D{|}_x$ is defined, $x$ and $s$ have the same type,\\ & and
                    $\sigma(x)$ is either undefined or equal to $s$ \\[\jot]
  \textsf{Backtrack} & $ ( \tuple{s}{m} \cdot \overline{t}{,}\; (s, D_0, \sigma_0) \cdot \overline{b}{,}\; D{,}\; \sigma)
                      \Pdtarrow
                      (s \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D_0{,}\; \sigma_0)$  \\
                      & if $D_0{|}_\zeta = D$ and $m = \textit{arity}(\zeta) - \textit{arity}(s)$ \\[\jot]
  \textsf{Success}    & $ ( [\,]{,}\; \overline{b}{,}\; D{,}\; \sigma ) \Pdtarrow (\mathit{val}(D){,}\; \sigma)$ \\
                      & if~$\mathit{val}(D)$ is defined
\end{tabular}

Above, $\cdot$ denotes prepending an element or a list to a list.
%
Intuitively, \textsf{AdvanceF} and \textsf{AdvanceX} move deeper in
the tree, generalizing cases A and B above to \lfhol{} terms.
\textsf{Backtrack} can be used to return to a previous state.
\textsf{Success} extracts the term~$t$ and data~$d$ stored in
an accepting node.

\newcommand\BDTARROW[1]{\rlap{\ensuremath{\Pdtarrow_{#1}\;}}\phantom{\Pdtarrow_\textsf{AdvanceX}\;}}

The following derivation illustrates how to locate a generalization of
$\cst{g}\; (\cst{g}\;\cst{a}\;\cst{b})$ in the tree $D_2$:\pagebreak[2]
%
\begin{align*}
 & ([\cst{g} \; (\cst{g} \; \cst{a} \; \cst{b})]{,}\;  [\,]{,}\;  D{,}\;  \emptyset) \\[-1\jot]
 \BDTARROW{\textsf{AdvanceF}}&
 ([\cst{g} \; \cst{a} \; \cst{b}]{,}\;  [(\cst{g} \; (\cst{g} \; \cst{a} \; \cst{b}), D, \emptyset)]{,}\;  D{|}_{\cst{g}}{,}\;  \emptyset)\\[-1\jot]
 \BDTARROW{\textsf{AdvanceX}}&
 ([\cst{b}]{,}\;  [(\cst{g} \; \cst{a} \; \cst{b}, D{|}_{\cst{g}}, \emptyset), \ldots]{,}\;  D{|}_{\cst{g}.y}{,}\;  \{y \mapsto \cst{g} \; \cst{a} \}) \\[-1\jot]
 \BDTARROW{\textsf{AdvanceX}}&
 ([\,]{,}\;  [(\cst{b}, D{|}_{\cst{g}.y}, \{y \mapsto \cst{g} \; \cst{a} \}), \ldots]{,}\; D{|}_{\cst{g}.y.x}{,}\; \{ y \mapsto \cst{g} \; \cst{a}{,}\; x \mapsto \cst{b} \}) \\[-1\jot]
 \BDTARROW{\textsf{Success}}& ((\cst{g} \; (y \; x), d){,}\; \{ y \mapsto \cst{g} \; \cst{a}{,}\; x \mapsto \cst{b} \})
\end{align*}

Let ${\Pdtarrow_{\mathsf{Advance}}} = {\Pdtarrow_{\mathsf{AdvanceF}}} \cup
{\Pdtarrow_{\mathsf{AdvanceX}}}$.
%and ${\Pdtarrow_{\mathsf{NoBacktrack}}} = {\Pdtarrow_{\mathsf{Advance}}} \cup
%{\Pdtarrow_{\mathsf{Success}}}$.
It is easy to show that \textsf{Backtrack} undoes an \textsf{Advance}
transition: % and can be avoided when starting from an initial state.

\begin{lemma}\label{lemma:ehoh:backtrack-inverses-advance} If $\quadruple \Pdtarrow_{\mathsf{Advance}} \quadruple'$, then
  $\quadruple' \Pdtarrow_{\mathsf{Backtrack}} \quadruple$.
  \end{lemma}
\begin{proof}
For both \textsf{Advance} steps, we show that \textsf{Backtrack} restores the state
properly. If \textsf{AdvanceF} was applied, we have
%
\begin{align*}
  ( \cst{f} \> \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma )
& \Pdtarrow_{\mathsf{AdvanceF}}
  (\tuple{s}{m} \cdot \overline{t}{,}\;
      (\cst{f} \> \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_\cst{f}{,}\; \sigma)
\\[-\jot]
& \Pdtarrow_{\mathsf{Backtrack}}
  ( \overline{t'}{,}\; \overline{b}{,}\; D{,}\; \sigma )
\end{align*}
We must show that $\overline{t'} = \cst{f} \> \tuple{s}{m} \mathrel{\cdot} \overline{t}$.
Let $k = \textit{arity}(\cst{f})$ and $l = \textit{arity}(\cst{f} \> \tuple{s}{m})$.
By definition of $k$, we have $m = k - l$, as in \textsf{Backtrack}'s side
condition. Thus, $\overline{t'} = \cst{f} \> \tuple{s}{m} \mathrel{\cdot}
\overline{t}$. The other case is
\begin{align*}
  ( s \; \tuple{s}{m} \mathrel{\cdot} \overline{t}{,}\; \overline{b}{,}\; D{,}\; \sigma )
& \Pdtarrow_{\mathsf{AdvanceX}}
  (\tuple{s}{m} \cdot \overline{t}{,}\; (s \; \tuple{s}{m}, D, \sigma) \cdot \overline{b}{,}\; D{|}_x{,}\; \sigma')
\\[-\jot]
& \Pdtarrow_{\mathsf{Backtrack}}
  ( \overline{t'}{,}\; \overline{b}{,}\; D{,}\; \sigma )
\end{align*}
where $\sigma' = \sigma[x \mapsto s]$.
Again, we must show that $\overline{t'} = s \; \tuple{s}{m} \mathrel{\cdot} \overline{t}$. Terms
$x$ and $s$ must have the same type for \textsf{AdvanceX} to be applicable; therefore, they have the same arity. Then, we conclude $m=
\textit{arity}(s) - \textit{arity}(s \; \tuple{s}{m} ) =  \textit{arity}(x) - \textit{arity}(s \; \tuple{s}{m} )$, as in \textsf{Backtrack}'s
side condition. Thus, $\overline{t'} = s \; \tuple{s}{m} \mathrel{\cdot} \overline{t}$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:advance-inverses-backtrack}
  If $\quadruple \Pdtarrow_{\mathsf{Advance}} \quadruple'
  \Pdtarrow_{\mathsf{Backtrack}} \quadruple''$, then
  $\quadruple'' = \quadruple$.
  \end{lemma}
\begin{proof}
  By Lemma~\ref{lemma:ehoh:backtrack-inverses-advance}, $\quadruple'
  \Pdtarrow_{\mathsf{Backtrack}} \quadruple$. Furthermore, \textsf{Backtrack} is
  clearly functional. Thus, $\quadruple'' = \quadruple$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:backtrack-remove}
  Let $\quadruple = ([t], [\,], D, \emptyset)$. If $\quadruple
  \Pdtarrow^* \quadruple'$, then $\quadruple \Pdtarrow_{\mathsf{Advance}}^*\allowbreak
  \quadruple'$.
\end{lemma}

\begin{proof}
  \looseness=-1
  Let $\quadruple = \quadruple{}_{\,\,0} \Pdtarrow \cdots \Pdtarrow \quadruple{}_{\,\,n} = \quadruple'$.
  Let $i$ be the index of the first transition of the form $\quadruple{}_{\,\,i} \Pdtarrow_{\mathsf{Backtrack}}
  \quadruple{}_{\,\,i+1}$. Since $\quadruple{}_{\,\,0}$'s backtracking stack is empty, we must
  have $i \not= 0$. Hence, we have $\quadruple{}_{\,\,i-1} \Pdtarrow_{\mathsf{Advance}}
  \quadruple{}_{\,\,i} \Pdtarrow_{\mathsf{Backtrack}} \quadruple{}_{\,\,i+1}$. By
  Lemma~\ref{lemma:ehoh:advance-inverses-backtrack}, $\quadruple{}_{\,\,i-1} =
  \quadruple{}_{\,\,i+1}$. Thus, we can shorten the derivation to $\quadruple{}_{\,\,0}
  \Pdtarrow \cdots \Pdtarrow \quadruple{}_{\,\,i-1} = \quadruple{}_{\,\,i+1} \Pdtarrow
  \cdots \Pdtarrow \quadruple{}_{\,\,n}$,
  thereby eliminating one \textsf{Backtrack} transition. By repeating this
  process, we eliminate all \textsf{Backtrack} transitions.
\end{proof}

\begin{lemma}\label{lemma:ehoh:termination-pdt}
  There exist no infinite chains of the form
  $\quadruple{}_{\,\,0} \Pdtarrow_\mathsf{Advance}\allowbreak \quadruple{}_{\,\,1} \Pdtarrow_\mathsf{Advance}
  \cdots$.
\end{lemma}
\begin{proof}
  With each \textsf{Advance} transition, the height of the discrimination tree
  decreases by at least one.
\end{proof}

Perfect discrimination trees match a single term against a set of terms. To
prove them correct, we will connect them to the transition system
$\Matcharrow$ for matching (Sect.~\ref{sec:ehoh:unif-match}). This connection\pagebreak[2] will
help us show that whenever a discrimination tree stores a generalization of a
query term, this generalization can be found.
%
To express the refinement, we introduce an intermediate transition
system, $\Matchiiarrow$, that focuses on a single pair of terms (like
$\Matcharrow$) but that solves the constraints in a depth-first, left-to-right
fashion and builds the substitution incrementally (like $\Pdtarrow$).
Its initial states are of the form $([s \MATCH t], \emptyset)$. Its transitions
are as follows:

\noindent
\begin{tabular}{ll}
  \textsf{Decompose} & $ (\cst{f} \; \tuple{s}{m} \MATCH \cst{f} \; \tuple{t}{m} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow {} ( (s_1 \MATCH t_1, \ldots, s_m \MATCH t_m) \cdot \overline{c}, \sigma ) $ \\[\jot]
  \textsf{DecomposeX} & $(x \; \tuple{s}{{m}} \MATCH u \; \tuple{t}{{m}} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow {} ( (s_1 \MATCH t_1, \ldots, s_m \MATCH t_m) \cdot \overline{c}, \sigma[x \mapsto u] )$ \\
                    & if $x$ and $u$ have the same type and either $\sigma(x)$ is undefined or $\sigma(x)=u$ \\[\jot]
  \textsf{Success} & $ ([\,], \, \sigma) \Matchiiarrow  \sigma $ \\[\jot]
  \textsf{Clash}    & $ (\cst{f} \; \tuple{s}{m} \MATCH \cst{g} \; \tuple{t}{n} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot $ \\[\jot]
  \textsf{ClashTypeX}    & $(x \; \tuple{s}{{m}} \MATCH u \; \tuple{t}{{m}} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot$ \\
                         & if~$x$ and $u$ have different types  \\[\jot]
  \textsf{ClashLenXF}    & $(x \; \tuple{s}{{m}} \MATCH \cst{f} \; \tuple{t}{{n}} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot$ \\
                         & if~$m > n$  \\[\jot]
  \textsf{ClashLenXY}    & $(x \; \tuple{s}{{m}} \MATCH y \; \tuple{t}{{n}} \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot$ \\
                         & if~$x \neq y$ and $m > n$ \\[\jot]
  \textsf{ClashFX}    & $(\cst{f} \; \overline{s} \MATCH x \; \overline{t}  \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot$ \\[\jot]
  \textsf{Double}    & $(x \; \tuple{s}{{m}} \MATCH u \; \tuple{t}{{m}}  \mathrel{\cdot} \overline{c}, \, \sigma) \Matchiiarrow \bot$ \\
                     & if $x$ and $u$ have the same type, $\sigma(x)$ is defined, and $\sigma(x) \neq u$ \\[\jot]
\end{tabular}

\newcommand\AlphaMatch[1]{\alpha(#1)}

We need an auxiliary function to convert $\Matchiiarrow$ states to
$\Matcharrow$ states.
Let
$\AlphaMatch{\{ x_1 \mapsto s_1,\allowbreak \ldots, x_m \mapsto s_m \}}
 = \{ x_1 \MATCH s_1,\ldots,\,\allowbreak x_m \MATCH s_m \}$,
$\AlphaMatch{\overline{c}, \sigma} =
\{ c \mid c \in \overline{c} \} \mathrel\cup \AlphaMatch{\sigma}$, and
$\AlphaMatch{\bot} = \bot$.
%
Moreover, let $\mathcal{S}$ range over states of the form $(\overline{c}, \sigma)$ and
$\mathcalx{R}$ additionally range over special states of the form $\sigma$ or
$\bot$.

\begin{lemma}\label{lemma:ehoh:match-refinement}
  If $\mathcal{S} \Matchiiarrow \mathcalx{R}$,
  then $\AlphaMatch{\mathcal{S}} \Matcharrow^* \AlphaMatch{\mathcalx{R}}$.
\end{lemma}
\begin{proof}
By case distinction on $\mathcalx{R}$. Let $\mathcal{S} = (\overline{c}, \sigma)$.

\medskip

\begin{sloppypar}%
  \noindent
  \textsc{Case} $\mathcalx{R} = (\overline{c'}, \sigma')$:\enskip
  Only $\Matchiiarrow_{\mathsf{Decompose}}$ and
  $\Matchiiarrow_{\mathsf{DecomposeX}}$ are possible.
  If $\Matchiiarrow_{\mathsf{Decompose}}$ is applied, then $\Matcharrow_{\mathsf{Decompose}}$
  is applicable and results in $\AlphaMatch{\mathcalx{R}}$.
  If $\Matchiiarrow_{\mathsf{DecomposeX}}$ is applied,
  we have either $m>0$, and $\Matcharrow_{\mathsf{DecomposeX}}$ is applicable,
  or $m = 0$, and $\AlphaMatch{\overline{c'},\sigma'} = \AlphaMatch{\mathcal{S}}$,
  which implies that the two states are connected by an idle transition of
  $\Matcharrow^*$.
  \end{sloppypar}

  \medskip

  \noindent
  \textsc{Case} $\mathcalx{R} = \bot$:\enskip
  All the $\Matchiiarrow$ rules resulting in $\bot$ except for \textsf{Double}
  have the same side conditions as the corresponding $\Matcharrow$ rules.
  $\Matchiiarrow_{\mathsf{Double}}$ corresponds to
  $\Matcharrow_{\mathsf{Double}}$ if $m=0$. If $m \neq 0$, we need
  an intermediate $\Matcharrow_{\mathsf{DecomposeX}}$ step before
  $\Matcharrow_{\mathsf{Double}}$ can be applied to derive $\bot$. Since
  $\Matchiiarrow_{\mathsf{Double}}$ is applicable, $\sigma(x) = u'
  \neq u$. Hence, $x \MATCH u'$ must be present in $\AlphaMatch{\overline{c},
  \sigma}$. $\Matcharrow_{\mathsf{DecomposeX}}$ will augment this set with
  $x \MATCH u$, enabling $\Matcharrow_{\mathsf{Double}}$.

  \medskip

  \noindent
  \textsc{Case} $\mathcalx{R} = \sigma$:\enskip
  The only possible rule is $\Matchiiarrow_{\mathsf{Success}}$,
  with $\overline{c}=[\,]$. Since $\AlphaMatch{\mathcal{S}} = \AlphaMatch{\sigma}$,
  this transition corresponds to an idle transition of $\Matcharrow^*$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:refined-match-partial-correctness}\,%
  If $\mathcal{S} \Matchiiarrow^{!} \mathcalx{R}$, then $\mathcalx{R}$ is either
  some substitution $\sigma'$ or $\bot$.
  If $\mathcal{S} \Matchiiarrow^{!} \sigma'$, then $\sigma'$ is the MGG
  of~$\AlphaMatch{\mathcal{S}}$.
  If $\mathcal{S} \Matchiiarrow^{!} \bot$, then $\AlphaMatch{\mathcal{S}}$ has
  no solutions.
  \end{lemma}
\begin{proof}
  %Let $\mathcal{S} = (\overline{c}, \sigma)$.

  First, we show that states $\mathcal{S}' = (\overline{c'}, \sigma')$ cannot be
  normal forms, by exhibiting transitions from such states. If $\overline{c'} =
  [\,]$, the $\Matchiiarrow_{\mathsf{Success}}$ rule would apply. Otherwise, let
  $\overline{c'} = c_1 \cdot \overline{c''}$ and consider the matching problem
  $\{c_1\} \cup \AlphaMatch{\sigma'}$. If this problem is in solved form, $c_1$
  is a constraint corresponding to a solved variable, and we can apply
  $\Matchiiarrow_{\mathsf{DecomposeX}}$ to move the constraint into the
  substitution. Otherwise, some $\Matcharrow$ rule can be applied. It
  necessarily focuses on $c_1$, since the constraints from
  $\AlphaMatch{\sigma'}$ correspond to solved variables. In all cases except for
  $\Matcharrow_{\mathsf{DecomposeX}}$, a homologous
  $\Matchiiarrow$ rule  can be applied to $\mathcal{S}'$. If $\Matcharrow_{\mathsf{DecomposeX}}$
  would make $\Matcharrow_{\mathsf{Double}}$ applicable, then we can apply $\Matchiiarrow_{\mathsf{Double}}$ to $\mathcal{S}'$;
  otherwise, $\Matchiiarrow_{\mathsf{DecomposeX}}$ is applicable.

  Second, by Lemma \ref{lemma:ehoh:match-refinement},
  if $\mathcal{S} \Matchiiarrow^{!} \sigma'$, then
  $\AlphaMatch{\mathcal{S}} \Matcharrow^* \AlphaMatch{\sigma'}$.
  By construction, $\AlphaMatch{\sigma'}$ is in solved form.
  Therefore, $\AlphaMatch{\mathcal{S}} \Matcharrow^{!} \AlphaMatch{\sigma'}$.
  By completeness of $\Matcharrow$, the substitution corresponding to
  $\AlphaMatch{\sigma'}$---that is, $\sigma'$---is the MGG of~$\AlphaMatch{\mathcal{S}}$.

  Third, by Lemma~\ref{lemma:ehoh:match-refinement}, if $\mathcal{S}
  \Matchiiarrow^{!} \bot$, then $\AlphaMatch{\mathcal{S}} \Matcharrow^{!} \bot$.
  By soundness of $\Matcharrow$, $\AlphaMatch{\mathcal{S}}$ has
  no solutions.
\end{proof}

\begin{lemma}\label{lemma:ehoh:refined-match-termination}\,%
  The relation $\Matchiiarrow$ is well founded.
  \end{lemma}
\begin{proof}
  By Lemma \ref{lemma:ehoh:match-refinement}, every $\Matchiiarrow$ transition
  corresponds to zero or more $\Matcharrow$ transitions.
  Since $\Matcharrow$ is well founded,
  the only transitions that can violate well-foundedness of $\Matchiiarrow$ are
  the ones that take idle $\Matcharrow^*$ transitions:
  $\Matchiiarrow_{\mathsf{DecomposeX}}$ for $m=0$ and
  $\Matchiiarrow_{\mathsf{Success}}$. The latter is terminal, so it cannot
  contribute to infinite chains.
  As for $\Matchiiarrow_{\mathsf{DecomposeX}}$, with $m=0$,
  it decreases the following measure~$\mu$, which the other rules nonstrictly
  decrease, with respect to the multiset extension of $<$ on natural numbers:
  $\mu([s_1 \MATCH t_1, \ldots, s_m \MATCH t_m], \sigma) =
    \{\left|s_1\right|, \ldots \left|s_m\right|\}$,
  where $\left|s\right|$ denotes the syntactic size of $s$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:refined-matching-solution}
  If term $s$ generalizes $t$, then $([s \MATCH t], \emptyset) \Matchiiarrow^! \sigma$,
  where $\sigma$ is the MGG of $s \MATCH t$.
\end{lemma}
\begin{proof}
By Lemma \ref{lemma:ehoh:refined-match-termination}, there exists a normal form
$\mathcalx{R}$ starting from $\mathcal{S} = ([s \MATCH t],\allowbreak \emptyset)$. Since $s
\MATCH t$ is solvable, by Lemma~\ref{lemma:ehoh:refined-match-partial-correctness},
and soundness of $\Matchiiarrow$ (a consequence of Lemma \ref{lemma:ehoh:match-refinement}
and soundness of $\Matcharrow$),
$\mathcalx{R}$ must be the MGG for $s$ and~$t$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:pdt-complete-help}
  If there exists a term $s \in \Term(D)$ that generalizes the query term~$t$,
  then there exists a derivation $([t], [\,],\allowbreak D,\allowbreak \emptyset) \Pdtarrow^! ((s, d),
  \sigma)$. %, where $\sigma$ is the MGG for $s$ and $t$.
  \end{lemma}
  \begin{proof}
    By Lemma \ref{lemma:ehoh:refined-matching-solution}, we know that $(s \MATCH t,
    \emptyset) \Matchiiarrow^! \sigma$ for each $s \in \Term(D)$ generalizing $t$.
    This means that there exists a derivation
    $([s \MATCH t], \emptyset) = (\overline{c_0}, \sigma_0) \Matchiiarrow \cdots \Matchiiarrow (\overline{c_n}, \sigma_n)\allowbreak \Matchiiarrow \sigma$.
    The $n$ first transitions must be \textsf{Decompose} or \textsf{DecomposeX},
    and the last transition must be \textsf{Success}.

    We show that there exists a derivation of the form
    $([t], [\,],\allowbreak D, \emptyset) = \quadruple{}_{\,\,0}\allowbreak \Pdtarrow \cdots
    \Pdtarrow \quadruple{}_{\,\,n} \Pdtarrow ((s, d),\sigma)$, where $\quadruple{}_{\,\,i} =
    (\overline{t_i}, \overline{b_i}, \allowbreak D_i, \sigma_i)$ for each $i$.
  %
    We define $\overline{t_i}$, $\overline{b_i}$, and $D_i$ as follows, for $i > 0$.
    The list $\overline{t_i}$ consists of the right-hand sides of the
    constraints~$\overline{c_i}$, in the same order.
    %todo:
    Let $\mathit{hd}$ be the function that extracts the head of a list. We set
          $b_i = (\mathit{hd}(\overline{t_{i-1}}),\allowbreak D_{i-1}, \sigma_{i-1})$. We know that $\overline{c_{i-1}}$ is
          nonempty, since there exists a transition $(\overline{c_{i-1}}, \sigma_{i-1}) \Matchiiarrow (\overline{c_i}, \sigma_i)$%
          ; thus, $\overline{t_{i-1}}$ is nonempty.
    If an accepting node storing~$s$ was reached in $n$~steps, the
      serialization of $s$ must be of the form $\zeta_1.\cdots.\zeta_n$. Take $D_i
      = D_{i-1}{|}_{\zeta_i}$.

    The sequence of states $\quadruple{}_{\,\,i}$ forms a
    derivation:
    If $(\overline{c_i},\allowbreak \sigma_i)\allowbreak \Matchiiarrow_\textsf{Decompose}
    (\tuple{c}{{i+1}},\allowbreak \sigma_{i+1})$, then $\quadruple{}_{\,\,i}
    \Pdtarrow_\textsf{AdvanceF} \allowbreak \quadruple{}_{\,\,i+1}$.
    If $(\overline{c_i},\allowbreak \sigma_i) \Matchiiarrow_\textsf{DecomposeX}
    (\tuple{c}{{i+1}},\allowbreak \sigma_{i+1})$, then $\quadruple{}_{\,\,i}
    \Pdtarrow_\textsf{AdvanceX} \quadruple{}_{\,\,i+1}$.
    If $(\overline{c_n},\allowbreak \sigma_n) \Matchiiarrow_\textsf{Success} \sigma$, then
    $\quadruple{}_{\,\,n} \Pdtarrow_\textsf{Success} ((s, d),\allowbreak \sigma)$.
\end{proof}

\begin{lemma}\label{lemma:ehoh:pdt-soundi}
  If $([t], [\,], D, \emptyset) \Pdtarrow^+ ((s, d), \sigma)$, then $s \in \Term(D)$
  and $\sigma$ is the MGG of $s \MATCH t$.
  \end{lemma}

  \begin{proof}
  Let $([t], [\,], D, \emptyset) =
  \quadruple{}_{\,\,0} \Pdtarrow \cdots \Pdtarrow \quadruple{}_{\,\,n}
  \Pdtarrow ((s, d), \sigma)$ be a derivation, where $\quadruple{}_{\,\,i} =
  (\overline{t_i}, \overline{b_i},\allowbreak D_i, \sigma_i)$ for each $i$. Without loss of
  generality, by Lemma \ref{lemma:ehoh:backtrack-remove}, we can assume that the
  derivation contains no \textsf{Backtrack} transitions.

  The first conjunct, $s \in \Term(D)$, clearly holds
  %by definition --- some invariant must be needed here
  for any term found
  from an initial state. To prove the second conjunct, we first introduce a
  function $\mathit{preord}$ that defines the preorder decomposition of a list
  of terms:
  $\mathit{preord}([\,]) = [\,]$ and
  $\mathit{preord}(\zeta \; \tuple{s}{n} \cdot \overline{xs}) =
  (\zeta{,}\; \tuple{s}{n} \cdot \overline{xs}) \cdot \mathit{preord}(\tuple{s}{n} \cdot \overline{xs})$.
  %
  Given a term~$s$, $\mathit{preord}([s])$ gives a sequence
  $(\zeta_1, \tuple{u}{1}), \dots, \allowbreak
  (\zeta_n, \tuple{u}{n})$.
  Since $s \in \Term(D)$,
  the sequence $D_0, \ldots, D_n$ follows the preorder serialization of $s$:
  $D_i = D_{i-1}|_{\zeta_i}$ for $i > 0$.

  Next, we show that there exists a derivation of the form
  $([s \MATCH t], \emptyset) = \mathcal{S}_0 \Matchiiarrow
  \cdots \Matchiiarrow \mathcal{S}_n \Matchiiarrow \sigma$, where
  $\mathcal{S}_i = (\overline{c_i}, \sigma_i)$. We define $\overline{c_i}$, for $i>0$,
  as the list of constraints whose left-hand sides are the elements of $\overline{u_i}$
  and right-hand sides are the elements of $\overline{t_i}$, in the order they appear in the respective lists.
  By inspecting the definition of $\mathit{preord}$
  and the changes each \textsf{Advance} step makes
  to the head of $\overline{t_i}$, we can see that $\overline{u_i}$ and $\overline{t_i}$  have the same length.
  The sequence of states $\mathcal{S}_i$ forms a derivation:
  %
  If $\quadruple{}_{\,\,i} \Pdtarrow_\textsf{AdvanceF} \quadruple{}_{\,\,i+1}$, then $\mathcal{S}_i \Matchiiarrow_\textsf{Decompose} \mathcal{S}_{i+1}$.
  If $\quadruple{}_{\,\,i} \Pdtarrow_\textsf{AdvanceX} \quadruple{}_{\,\,i+1}$, then $\mathcal{S}_i\allowbreak \Matchiiarrow_\textsf{DecomposeX} \mathcal{S}_{i+1}$.
  If $\quadruple{}_{\,\,n} \Pdtarrow_\textsf{Success} \sigma$, then
  $\mathcal{S}_n\allowbreak \Matchiiarrow_\textsf{Success} \sigma$.
\end{proof}

\begin{theorem}[Total Correctness]\label{theorem:ehoh:pdt-complete}\,%
  Let $D$ be a perfect discrimination tree and $t$ be a term.
  The sets $\{ s \in \Term(D) \mid \exists \sigma.\; \sigma(s) = t \}$
  and $\{ s \mid \exists d, \sigma.\; ([t], [\,], D, \emptyset) \allowbreak
  \Pdtarrow^! ((s, d), \sigma) \}$ are equal.
\end{theorem}
\begin{proof}
  This follows from Lemmas \ref{lemma:ehoh:pdt-complete-help} and \ref{lemma:ehoh:pdt-soundi}.
\end{proof}

The theorem tells us that given a term~$t$, all generalizations $s$ stored
in the perfect discrimination tree can be found, but it does not exclude
nondeterminism. Often, both \textsf{AdvanceF} and \textsf{AdvanceX} are
applicable. To find all generalizations, we need to follow both transitions.
But for some applications, it is enough to find a single generalization.

To cater for both types of applications, E provides iterators that store the
state of a traversal. After an iterator is initialized
with the root node $D$ and the query term $t$, each call to
\textsc{FindNextVal} will move the iterator to the next node that generalizes
the query term and stores a value, indicating an accepting node. After all
such nodes have been traversed, the iterator is set to point to
$\mathit{Null}$.

The following definitions constitute the high-level interface for
iterating through values incrementally or for obtaining all values of nodes that
store generalizations of the query term in $D$.

\begin{quotex}
  \MyFunction{InitIter}{PDTNode $D$, Term $t$}
  \q $i \gets \textsc{Iterator}()$ \\
  \q $(i.\mathit{node}, i.\mathit{t\_stack}, i.\mathit{t\_proc}, i.\mathit{c\_iter}) \gets (D, [t], [\,], \textit{Start})$ \\
  \q \MyReturn $i$

  \vskip 1.5\jot

  \MyProcedure{FindNextVal}{Iterator $i$}
  \q \MyDo
  \qq \textsc{FindNextNode}($i$) \\
  \q \MyWhileOfDo{$i.\mathit{node} \not= \mathit{Null} \mathrel\land
    \\ \qqq  (\lnot\, i.\mathit{t\_stack}.\mathit{isEmpty}()
    \mathrel\lor \lnot\, i.\mathit{node}.\mathit{has\_val}())$}

  \vskip 1.5\jot

  \MyFunction{AllVals}{PDTNode $D$, Term $t$}
  \q $\mathit{i} \gets \textsc{InitIter}(D, t)$ \\
  \q $\textsc{FindNextVal}(i)$ \\
  \q $\mathit{res} \gets \emptyset$ \\
  \q \MyWhile{$\mathit{i.node} \neq \mathit{Null}$}
  \qq $\mathit{res} \gets \mathit{res} \cup \{ \mathit{i.node.val()} \}$ \\
  \qq $\textsc{FindNextVal}(i)$
  \\[\jot]
  \q \MyReturn $\mathit{res}$
  \end{quotex}

  The core functionality is implemented in \textsc{FindNext\-Node}, presented below. This
  procedure moves the iterator to the next node that has not been explored in the
  search for generalization, or \textit{Null} if the entire tree has been
  traversed. It first goes through all child nodes labeled with a variable before
  possibly visiting the child node labeled with a function symbol. We assume that
  we can iterate through the children of a node using a function
  \textsc{NextVarChild} that, given a tree node and iterator through children,
  advances the iterator to the child representing the next variable. Furthermore, we
  assume that the iterator can also be in the distinguished states \textit{Start}
  and \textit{End}.
  \textit{Start} indicates that no child has been visited yet; \textit{End}
  indicates that we have visited all children. Finally, the expression
  $n.\mathit{child}(\zeta)$ returns a child of the node $n$ labeled $\zeta$ if
  such a child exists or $\mathit{Null}$ otherwise.

  \begin{quotex}
  \MyProcedure{FindNextNode}{Iterator $i$}
  \q \MyIf{$i.\mathit{t\_stack.isEmpty}()$}
  \qq \textsc{BacktrackToVar}($i$)
  \\[\jot]
  \q $\mathit{advanced} \gets \mathit{False}$ \\
  \q \MyWhile{$i.\mathit{node} \neq \mathit{Null} \mathrel\land \lnot\, \mathit{advanced}$}
  %\qq \MyIf{$i.\mathit{c\_iter} < \textit{End}$}
  \qq \MyWhile{$i.\mathit{c\_iter} \neq \textit{End} \mathrel\land \lnot\, \mathit{advanced}$}
  \qqq $i.\mathit{c\_iter} \gets \textsc{NextVarChild}(i.\mathit{node}, i.\mathit{c\_iter})$ \\
  \qqq \MyIf{$i.\mathit{c\_iter} \neq \textit{End}$}
  \qqqq $x \gets i.\mathit{c\_iter}.\mathit{var}()$ \\
  \qqqq $t \gets i.\mathit{t\_stack}.\mathit{top}()$ \\
  \qqqq $\mathit{s} \gets \textsc{GobblePrefix}(x, t)$ \\
  \qqqq \MyIf{$\mathit{s} \neq \mathit{Null} \mathrel\land \\ \qqqqqq
                (x.\mathit{binding} = \mathit{Null} \mathrel\lor x.\mathit{binding} = \mathit{s})$}
  \qqqqq $i.\mathit{t\_stack}.\mathit{pop}()$ \\
  \qqqqq \MyForDownto{$j \gets t.\mathit{num\_args}$\\\qqqqqqq\ignorespaces}{$\mathit{s}.\mathit{num\_args}+1$}
  \qqqqqq $i.\mathit{t\_stack}.\mathit{push}(t.\mathit{args}[j])$ \\
  \qqqqq \MyIf{$x.\mathit{binding} = \mathit{Null}$}
  \qqqqqq $x.\mathit{binding} \gets \mathit{s}$ \\
  \qqqqqq $i.\mathit{t\_proc}.\mathit{push}
                      ( (t, i.\mathit{node}, i.\mathit{c\_iter}, \mathit{True} ) )$ \\
  \qqqqq \MyElse
  \qqqqqq $i.\mathit{t\_proc}.\mathit{push}
                      ( (t, i.\mathit{node}, i.\mathit{c\_iter}, \mathit{False}) )$
  \\[\jot]
  \qqqqq $ \mathit{i.\mathit{node}} \gets i.\mathit{node}.child(x)$ \\
  \qqqqq $ \mathit{advanced} \gets \mathit{True}$
  \\[\jot]
  \qq $t \gets i.\mathit{t\_stack}.\mathit{top}()$ \\
  \qq \MyIf{$i.\mathit{c\_iter} = \textit{End} \mathrel\land
          \lnot\, t.\mathit{head}.\mathit{isVar}()  \\ \qqq \mathrel\land
          D.\mathit{child}(t.\mathit{head}) \neq \mathit{Null}$}
  \qqq $i.\mathit{t\_stack}.\mathit{pop}()$ \\
  \qqq \MyForDownto{$j \gets t.\mathit{num\_args}$}{$1$}
  \qqqq $i.\mathit{t\_stack}.\mathit{push}(t.\mathit{args}[j])$ \\
  \qqq $i.\mathit{t\_proc}.\mathit{push}
              ( (t, i.\mathit{node}, \mathit{End}, \mathit{False}) )$ \\
  \qqq $\mathit{i.\mathit{node}} \gets i.\mathit{node}.\mathit{child}(t.\mathit{head})$ \\
  \qqq $\mathit{advanced} \gets \mathit{True}$
  \\[\jot]
  \qq \MyIf{$\lnot\, \mathit{advanced}$}
  \qqq \textsc{BacktrackToVar}$(i)$ \\
  \qq \MyElse
  \qqq $i.\mathit{c\_iter} \gets \textit{Start}$

  \vskip 2\jot

  \MyProcedure{BacktrackToVar}{Iterator $i$}
  \q \MyForever
  \qq \MyIf{$i.\mathit{t\_proc}.\mathit{isEmpty}()$}
  \qqq $i.\mathit{node} \gets \mathit{Null}$ \\
  \qqq \MyReturn \\
  \qq \MyElse
  \qqq $(t, D, \mathit{c\_iter}, \mathit{var\_unbound}) \gets i.\mathit{t\_proc}.\mathit{pop}()$ \\
  \qqq $\mathit{label\_arity} \gets i.\mathit{node}.\mathit{label}.\mathit{type}.\mathit{arity}$ \\
  \qqq $\mathit{t\_arity} \gets t.\mathit{type}.\mathit{arity}$
  \\[\jot]
  \qqq \MyForTo{$i \gets 1$}{$\mathit{label\_arity} - \mathit{t\_arity}$}
  \qqqq $i.\mathit{t\_stack}.\mathit{pop}()$
  \\[\jot]
  \qqq $i.\mathit{t\_stack}.\mathit{push}(t)$ \\
  \qqq $i.\mathit{node} \gets D$ \\
  \qqq $i.\mathit{c\_iter} \gets \mathit{c\_iter}$
  \\[\jot]
  \qqq \MyIf{$\mathit{var\_unbound}$}
  \qqqq $i.\mathit{node}.\mathit{label}.\mathit{binding} \gets \mathit{Null}$
  \\[\jot]
  \qqq \MyIf{$\mathit{c\_iter} \neq\textit{End}$}
  \qqqq \MyReturn
  \end{quotex}

The pseudocode uses a slightly different representation of backtracking tuples
than ${\Pdtarrow}$. In the \textsf{AdvanceX} rule,
$\sigma$ changes only if the variable~$x$ was previously not bound.
Instead of creating and storing substitutions explicitly, we simply remember
whether the variable was bound in this step or not, in the
$\mathit{var\_unbound}$ tuple component. Then we rely on the label~$x$ of the
current node and its \textit{binding} field to carry the substitutions.
Similarly, since our strategy is to traverse the tree by first visiting the
variable-labeled child nodes, we need to remember how far we have come with
this traversal. We store this information in the $\mathit{c\_iter}$ tuple
component.

\subsection{Fingerprint Indices}
\looseness=-1
Fingerprint indices \cite{ss-12-fp-indexing} trade perfect indexing for
a compact memory representation and more flexible retrieval conditions.
The basic idea is to compare terms by looking only at a few
predefined sample positions. If we know that term $s$ has
symbol $\cst{f}$ at the head of the subterm at~2.1 and term $t$ has
$\cst{g}$ at the same position, we can immediately conclude that $s$ and $t$
are not unifiable.

% They are
% based on the idea that when one samples a term position only the following
% situations may happen: position corresponds to
% a variable or a complex term; position is invalid---but under some substitution it
% can become valid; position is invalid and no substitution makes it valid.

Let $\AA$ (``at a variable''),
$\BB$ (``below a variable''), and
$\NN$ (``nonexistent'') be distinguished symbols\begin{rep}
not present in the signature, and let $q < p$ denote that position $q$ is a proper prefix of $p$\end{rep}
(e.g., $\varepsilon < 2 < 2.1$). Given a term $t$ and a position~$p$, the
\emph{fingerprint function} $\Gfpf$ is defined as
\[
  \Gfpf(t,p) =
  \begin{cases}
    \cst{f} & \text{if $t|_p$ has a symbol head $\cst{f}$} \\[-\jot]
    \AA & \text{if $t|_p$ is a variable} \\[-\jot]
    \BB & \text{if $t|_q$ is a variable for some $q < p$} \\[-\jot]
    \NN & \text{otherwise}
  \end{cases}
\]

Based on a fixed tuple of positions $\tuple{p}{n}$,
the \emph{fingerprint} of a term $t$ is defined as
$\mathit\Fp(t) = \bigl(\Gfpf(t, p_1), \ldots, \Gfpf(t, p_n)\bigr).$ To compare
two terms $s$ and $t$, it suffices to check that their fingerprints are
componentwise compatible using the following unification and matching
matrices, respectively:
%
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\confrep{
\begin{align*}
& \hbox{\begin{tabular}{@{}C{0.75em}|C{0.75em}C{0.75em}C{0.75em}C{0.75em}C{0.75em}@{}}
& $\cst{f}_1$ & $\cst{f}_2$ & \AA & \BB & \NN \\
\hline
$\cst{f}_1\vphantom{^{()}}$ & & \xmark & & & \xmark \\
\AA & & & & & \xmark \\
\BB & & & & & \\
\NN & \xmark & \xmark & \xmark & &
\end{tabular}}
&
& \hbox{\begin{tabular}{@{}C{0.75em}|C{0.75em}C{0.75em}C{0.75em}C{0.75em}C{0.75em}@{}}
& $\cst{f}_1$ & $\cst{f}_2$ & \AA & \BB & \NN \\
\hline
$\cst{f}_1\vphantom{^{()}}$ & & \xmark & \xmark & \xmark & \xmark \\
\AA & & & & \xmark & \xmark \\
\BB & & & & & \\
\NN & \xmark & \xmark & \xmark & \xmark &
\end{tabular}}
\end{align*}
}{
\begin{align*}
& \hbox{\begin{tabular}{@{}C{0.75em}|C{0.75em}C{0.75em}C{0.75em}C{0.75em}C{0.75em}@{}}
& $\cst{f}_1$ & $\cst{f}_2$ & \AA & \BB & \NN \\
\hline
$\cst{f}_1\vphantom{^{()}}$ & & \xmark & & & \xmark \\
$\cst{f}_2$ & \xmark & & & & \xmark \\
\AA & & & & & \xmark \\
\BB & & & & & \\
\NN & \xmark & \xmark & \xmark & &
\end{tabular}}
&
& \hbox{\begin{tabular}{@{}C{0.75em}|C{0.75em}C{0.75em}C{0.75em}C{0.75em}C{0.75em}@{}}
& $\cst{f}_1$ & $\cst{f}_2$ & \AA & \BB & \NN \\
\hline
$\cst{f}_1\vphantom{^{()}}$ & & \xmark & \xmark & \xmark & \xmark \\
$\cst{f}_2$ & \xmark & & \xmark & \xmark & \xmark \\
\AA & & & & \xmark & \xmark \\
\BB & & & & & \\
\NN & \xmark & \xmark & \xmark & \xmark &
\end{tabular}}
\end{align*}
}%
%
The rows and columns correspond to $s$ and $t$, respectively. The
metavariables $\cst{f}_1$~and~$\cst{f}_2$ represent arbitrary distinct symbols.
Incompatibility is indicated by $\xmark{}$.

\newcommand\fpok{\text{--}}

As an example,
let $(\varepsilon, 1,\allowbreak 2,\allowbreak 1.1,\allowbreak 1.2,\allowbreak 2.1, 2.2)$
be the sample positions,
and let $s = \cst{f}(\cst{a}, x)$ and $t = \cst{f}(\cst{g}(x), \cst{g}(\cst{a}))$
be the terms to unify. Their fingerprints are
$\Fp(s) = (\cst{f}, \cst{a}, \AA, \NN, \NN,\allowbreak \BB, \BB)$
and
$\Fp(t) = (\cst{f}, \cst{g}, \cst{g}, \AA, \NN, \cst{a}, \NN)$.
%
Using the left matrix, we compute the compatibility vector
%
$(\fpok, \xmark, \fpok, \xmark, \fpok, \fpok, \fpok).$
%
The mismatches at positions 1 and 1.1 indicate that $s$ and $t$ are not unifiable.

A fingerprint index is a trie that stores a term set $T$ keyed by fingerprint.
The term $\cst{f}(\cst{g}(x), \cst{g}(\cst{a}))$ above would
be stored in the node addressed by
$\cst{f}.\cst{g}.\cst{g}.\AA.\NN.\cst{a}.\NN$, together with other
terms that share the same fingerprint.
This scheme makes it possible to unify or match a query term $s$
against all the terms $T$ in one traversal. Once a node storing the terms $U
\subseteq T$ has been reached, due to overapproximation we must apply
unification or matching on $s$ and each $u \in U.$ % to obtain correct results.

When adapting this data structure to \lfhol{}, we must first choose a
suitable notion of position in a term. Conventionally, higher-order
positions are strings over $\{1, 2\}$, but this is not graceful.
% indicating, for each binary
% application $t_1\;t_2$, which term $t_i$ to follow.
Instead, it is preferable to generalize the first-order notion to
flattened \lfhol{} terms---e.g.,
$x\> \cst{a} \> \cst{b} \> |_{1} = \cst{a}$
and $x\> \cst{a} \> \cst{b} \> |_{2} = \cst{b}.$
%
However, this approach fails on applied variables. For example,
although $x \> \cst{b}$ and $\cst{f} \> \cst{a} \> \cst{b}$ are
unifiable (using $\{ x \mapsto \cst{f} \> \cst{a} \}$), sampling position 1
would yield a clash between
$\cst{b}$~and~$\cst{a}.$ To ensure that positions remain stable under
substitution, we propose to number arguments in reverse:
$t {|}^\varepsilon = t$ and
$\zeta \> t_n \, \ldots \, t_1  |^{i.p} = t_{i} |^p$
if $1 \le i \le n$.
We use a nonstandard notation, $t|^p$, for this nonstandard
notion. The operation is undefined for out-of-bound indices.

\begin{lemma}\label{lemma:fpindex-aux}
  Let $s$ and $t$ be unifiable terms, and let $p$ be a position such that the
  subterms $s|^{p}$ and $t|^{p}$ are defined. Then $s|^{p}$ and $t|^{p}$ are
  unifiable.
\end{lemma}
\begin{proof}
  By structural induction on $p$. The case $p = \varepsilon$ is trivial.

  \medskip

  \noindent
  \textsc{Case $p = q.i$}:\enskip
  Let $s|^{q} = \zeta \; s_m \, \ldots \, s_1$ and $t|^{q} = \eta \; t_n \, \ldots \, t_1$.
  Since $p$ is defined in both $s$ and $t$, we have
   $s|^{p} = s_i$ and $t|^{p} = t_i$.
  By the induction hypothesis, $s|^{q}$ and $t|^{q}$ are
  unifiable, meaning that there exists a substitution $\sigma$
  such that $\sigma(\zeta \; s_m \, \ldots \, s_1) = \sigma(\eta \; t_n \, \ldots \, t_1)$.
  Hence, $\sigma(s_1) = \sigma(t_1)$, \ldots, $\sigma(s_i) = \sigma(t_i)$---i.e.,
  $\sigma(s|^{p}) = \sigma(t|^{p})$.
\end{proof}

Let $t{\langle}{}^p$ denote the subterm $t{|}^q$ such that $q$ is the longest
prefix of $p$ for which $t{|}^q$ is defined.
%
The \lfhol{} version of the fingerprint function is defined as follows:
%
\[
  \GfpfRTL(t,p) =
  \begin{cases}
    \cst{f} & \text{if $t|^p$ has a symbol head $\cst{f}$} \\
    \AA & \text{if $t|^p$ has a variable head} \\
    \BB & \text{if $t|^p$ is undefined} \\[-\jot]
        & \text{but $t{\langle}{}^p$ has a variable head} \\
    \NN & \text{otherwise}
  \end{cases}
\]
%
Except for the reversed numbering scheme,
$\GfpfRTL$ coincides with $\Gfpf$ on first-order terms.
The fingerprint $\FpRTL(t)$ of a term $t$ is defined analogously as before,
and the same compatibility matrices can be used.

The key difference between $\Gfpf$ and $\GfpfRTL$ concerns applied variable.
Given the sample positions $(\varepsilon, 2, 1)$,
the fingerprint of $x$ is $(\AA, \BB, \BB)$ as before, whereas
the fingerprint of $x \; \cst{c}$ is $(\AA, \BB, \cst{c}).$
%
As another example,
let $(\varepsilon, 2,\allowbreak 1,\allowbreak 2.2,\allowbreak 2.1,\allowbreak 1.2, 1.1)$
be the sample positions,
and let $s = x \; (\cst{f} \; \cst{b} \; \cst{c})$ and $t = \cst{g}\; \cst{a}\; (y \; \cst{d}).$
Their fingerprints are
$\FpRTL(s) = (\AA, \BB, \cst{f}, \BB, \BB, \cst{b}, \cst{c})$ and
$\FpRTL(t) = (\cst{g}, \cst{a}, \AA, \NN, \NN, \BB, \cst{d})$.
%
The terms are not unifiable due to the incompatibility at position 1.1
($\cst{c}$ vs.~$\cst{d}$).

We can easily support prefix optimization for both terms $s$ and $t$ being
compared: We simply add enough fresh variables as arguments to ensure that $s$
and $t$ are fully applied before computing their fingerprints.

\begin{lemma}\label{lemma:ehoh:fpindex-gfpf-sound-unif}
  If terms $s$ and $t$ are unifiable, then $\GfpfRTL(s, p)$ and $\GfpfRTL(t,
  p)$ are compatible according to the unification matrix.
  If $s$ generalizes $t$, then $\GfpfRTL(s, p)$ and $\GfpfRTL(t, p)$ are
  compatible according to the matching matrix.
\end{lemma}
\begin{proof}
  We focus on the case of unification. By contraposition, it suffices to
  consider the eight blank cells in the unification matrix, where the rows
  correspond to $\GfpfRTL(s, p)$ and the columns correspond to $\GfpfRTL(t,
  p)$. Since unifiability is a symmetric relation, we can rule out four cases.

\medskip

\noindent
\textsc{Case} $\cst{f}_1$--$\cst{f}_2$:\enskip
By definition of $\GfpfRTL$, $s|^{p}$ and $t|^{p}$ must be of the forms
$\cst{f}_1 \; \overline{s}$ and $\cst{f}_2 \; \overline{t}$, respectively.
Clearly, $s|^{p}$ and $t|^{p}$ are not unifiable. By Lemma
\ref{lemma:fpindex-aux}, $s$ and $t$ are not unifiable.

\medskip

\noindent
\textsc{Case} $\cst{f}_1$--$\NN$, $\cst{f}_2$--$\NN$, \textsc{or} $\AA$--$\NN$:\enskip
From $\GfpfRTL(t, p) = \NN$, we deduce that
$p \neq \varepsilon$.
Let $p = q.i.r$, where $q$ is the longest prefix such that $\GfpfRTL(t, q) \neq \NN$.
Since $\GfpfRTL(t, q.i) = \NN$, the head of $t|^{q}$ must be some symbol $\cst{g}$.
(For a variable head, we would have $\GfpfRTL(t, q.i) = \BB$.)
Hence, $t|^{q}$ has the form $\cst{g} \; t_n \, \ldots \, t_1$, for $n < i$.
Since $q.i$ is a legal position in $s$, $s|^{q}$ has the form
$\zeta \; s_m \, \ldots \, s_1$, with $i \le m$. A necessary condition for
$\sigma(s|^{q}) = \sigma(t|^{q})$ is that
$\sigma(\zeta \; s_m \, \ldots \, s_{n+1}) = \sigma(\cst{g})$,
but this is impossible because the left-hand side is an application (since
$n < m$), whereas the right-hand side is the symbol $\cst{g}$.
By Lemma \ref{lemma:fpindex-aux}, $s$ and $t$ are not unifiable.
\end{proof}

\begin{corollary}[Overapproximation]\,%
  If $s$ and $t$ are unifiable terms, then $\FpRTL(s)$ and $\FpRTL(t)$ are
  compatible according to the unification matrix.
  If $s$ generalizes $t$, then $\FpRTL(s)$ and $\FpRTL(t)$ are
  compatible according to the matching matrix.
\end{corollary}
%\begin{proof}
%This follows from Lemma~\ref{lemma:ehoh:fpindex-gfpf-sound-unif}.\qed
%\end{proof}

% \subsection{Feature Vector Indices}

% A clause $C$
% subsumes a clause $D$ if there exists a substitution $\sigma$ such that
% $\sigma(C) \subseteq D.$
% Subsumption is a crucial operation to prune the search space.
%  Feature-vector indices
% \cite{ss-2013-feature-vector} are an imperfect indexing data structure
% that can be used to retrieve clauses that subsume a query clause
% % (forward subsumption)
% or that are subsumed by the query clause.\
% %
% Unlike for discrimination trees and fingerprint indices, no changes were
% necessary to adapt feature vectors indices to \lfhol{}. All the predefined
% features make sense in \lfhol{}. % and are compatible with subsumption.

\section{Inference Rules}
\label{sec:ehoh:inferences}

\section{Heuristics}
\label{sec:ehoh:heuristics}

\section{Preprocessing}
\label{sec:ehoh:preprocessing}

\section{Evaluation}
\label{sec:ehoh:evaluation}

\section{Discussion and Related Work}
\label{sec:ehoh:discussion-and-related-work}

\section{Conclusion}
\label{sec:ehoh:conclusion}