### 2022-03-04 bools/bools.tex
bools

\chapter{Boolean Reasoning in a Higher-Order Superposition Prover}
\setheader{Boolean Reasoning in a Higher-Order Superposition Prover}
\label{ch:bools}

\blfootnote{In this work I implemented and evalauted most of the described approaches.
Visa Nummelin implemented the FOOL preprocessing module.}

\authors{Joint work with Visa Nummelin}

\begin{abstract}
    We present a pragmatic approach to extending a Boolean-free higher-order
    superposition calculus to support Boolean reasoning. Our approach extends
    inference rules that have been used only in a first-order setting, uses some
    well-known rules previously implemented in higher-order provers, as well as new rules.
    We have implemented the approach in the Zipperposition
    theorem prover. The evaluation shows highly competitive performance of our approach
    and clear improvement over previous techniques.
\end{abstract}

\newpage

\section{Introduction} 
\label{sect:bool:introduction}

In the last decades, automatic theorem provers have been used
successfully as backends to ``hammers'' in proof assistants
\cite{ku-15-holyhammer, pb-12-sh} and to software
verifiers \cite{fp-13-why3}. Most advanced provers, such as CVC4
\cite{cbetal-11-cvc4}, E \cite{scv-19-e23}, and Vampire
\cite{lkav-13-vampire}, are based on first-order logic, whereas
most frontends that use them are based on versions of higher-order logic. Thus,
there is a large gap in expressiveness between front- and backends.
%
\looseness=-1
This gap is bridged using well-known translations from higher-order to
first-order logic \cite{ar-70-hol,
mp-08-trans}. However, as shown in Chapter \ref{ch:ehoh}, translations are usually less efficient than native support. The
distinguishing features of higher-order logic used by proof assistants that the
translation must eliminate include
$\lambda$-binders, function extensionality---the property that functions are
equal if they agree on every argument, described by the axiom $\iforall xy. \allowbreak\, (\iforall z. \, x \, z \ieq y \,z)
\iimplies x \ieq y$, and formulas occurring as arguments of function symbols \cite{mp-08-trans}.

\looseness=-1
A group of authors including Vukmirovi\'c\ \cite{bbtvw-21-sup-lam} designed an
extension of superposition for extensional Boolean-free higher-order logic. The extension removes the need to translate the
first two above mentioned features of higher-order logic. Kotelnikov et al.\
\cite{kotelnikov-15-fool,kotelnikov-16-fool} extended the language of
first-order logic to support the third feature of higher-order logic that
requires translation. They described two approaches: One based on calculus-level
treatment of Booleans and the other, which requires no changes to the calculus,
based on preprocessing.

\looseness=-1 To fully bridge the gap between higher-order and first-order tools,
we combine the two approaches: we use the efficient higher-order superposition
calculus and extend it with inference rules that reason with
Boolean terms. In early work, Kotelnikov et al.\ 
\cite{kotelnikov-15-fool} have described a \emph{FOOL paramodulation}
rule that, under some order requirements, removes the need for the axiom
describing the Boolean domain---$\iforall p. \; p \ieq \itrue \llor p
\ieq \ifalse$. In this approach, it is assumed that a problem with formulas occurring as
arguments of symbols is translated to first-order logic.

\looseness=-1
The backbone of our approach is based on an extension of this rule to higher-order
logic: We do not translate away any Boolean structure that is nested inside
non-Boolean terms and allow our rule to hoist the nested Booleans to the
literal level. Then, we clausify the resulting formula (i.e., a clause containing formulas in literals) using a new rule.

\looseness=-1 An important feature that we inherit by building on top of Bentkamp
et al. \cite{bbtvw-21-sup-lam} is support for (function) extensionality. Moving to higher-order
logic with Booleans also means that we need to consider \emph{Boolean extensionality}: $\iforall pq. \, (p
\iequiv q) \iimplies p \ieq q$. We extend the rules of Bentkamp et al.\ 
that treat function extensionality to also treat Boolean extensionality.

\looseness=-1
Rules that extend the two orthogonal approaches form the basis of our support
for Boolean reasoning (Section \ref{sect:bool:native}). In addition, we have
implemented rules that are inspired by the ones used in the
higher-order provers Leo-III \cite{sb-21-leo3} and Satallax
\cite{cb-12-satallax}, such as elimination of Leibniz equality, primitive
instantiation and treatment of choice operator \cite{pa-01-classical-ty-thy}. We
have also designed new rules that use higher-order unification to resolve
Boolean formulas that are hoisted to literal level, delay clausification of non-atomic
literals, reason about formulas under $\lambda$-binders, and many others. Even
though the rules we use are inspired by the ones of refutationally complete
higher-order provers, we do not guarantee completeness of our extension of
$\lambda$-superposition.

\looseness=-1
We compare our native approach with two alternatives based on
preprocessing (Section \ref{sect:bool:alternative}). First, we compare it to  an
axiomatization of the theory of Booleans. Second, inspired by work of
Kotelnikov et al.\ \cite{kotelnikov-16-fool}, we implemented the preprocessing approach that does not
require introduction of Boolean axioms.
We also discuss some examples, coming from TPTP \cite{gs-17-tptp}, that
illustrate advantages and disadvantages of our approach (Section \ref{sect:bool:examples}).

Our approach is implemented in the Zipperposition theorem prover
\cite{sc-15-simon-phd,sc-supind-17}. Zipperposition is an easily
extensible open source prover that Bentkamp et al.\ used to implement their
higher-order superposition calculus. We further extend their implementation.

\looseness=-1
We performed an extensive evaluation of our approach (Section \ref{sect:bool:eval}).
In addition to evaluating different configurations of our new rules, we have
compared them to full higher-order provers CVC4, Leo-III,  Satallax and Vampire.
The results suggest that it is beneficial to natively support Boolean reasoning
-- the approach outperforms preprocessing-based approaches. Furthermore, it is
very competitive with state-of-the-art higher order provers. We discuss the differences between our approach and the
approaches we base on, as well as related approaches (Section
\ref{sect:bool:discussion}).

\section{Background} 
\label{sect:bool:background}

We base our work (and parts of the following text) on Bentkamp et al.'s \cite{bbtvw-21-sup-lam} extensional
polymorphic clausal higher-order logic. We extend the syntax of this logic by
adding logical connectives to the signature. The semantic of the logic
is extended by interpreting Boolean type $o$ as a two-element domain. This
amounts to extending Bentkamp et al's fragment of higher-order logic to
full-higher order logic (HOL). Taking a different perspective, this logic is an extension of the one
presented in Chapter \ref{ch:ehoh} with $\lambda$ abstraction, polymorphic
types, and native logical connectives. For reference, we provide precise definition of this polymorphic logic.

\looseness=-1
A signature is a quadruple $(\Sigmaty, \Vty, \Sigma, \VV)$ where $\Sigmaty$ is a
set of type constructors, $\Vty$ is a set of type variables and $\Sigma$ and
$\VV$ are sets of constants and term variables, respectively. We require nullary
type constructor $o$, as well as binary constructor $\rightarrow$
to be in $\Sigmaty$. A type $\tau, \upsilon$ is either a type variable $\alpha \in
\Vty$ or of the form $\kappa(\tau_1, \ldots \tau_n)$ where $\kappa$ is an
$n$-ary type constructor. We write $\kappa$ for $\kappa()$, $\tau \rightarrow
\upsilon$ for $\rightarrow(\tau, \upsilon)$, and we drop parentheses to shorten 
$\tau_1 \rightarrow (\cdots \rightarrow (\tau_{n-1} \rightarrow \tau_n) \cdots)$ into $\tau_1 \rightarrow \cdots \rightarrow
\tau_n$. Each symbol in $\Sigma$ is
assigned a type declaration of the form $\forallty{\tuplen{\alpha}} \tau$ where all variables
occurring in $\tau$ are among $\tuplen{\alpha}$.

\looseness=-1
Function symbols $\cst{a}, \cst{b}, \cst{f}, \cst{g}, \ldots$ are elements of
$\Sigma$; their type declarations are written as $\cst{f} :
\forallty{\tuplen{\alpha}} \tau$. Set $\VV$ contains both free and bound variables, following
the distinction introduced in Sect.~\ref{sec:pre:hol}. Term variables from the set $\VV$ are written
$F,G,X,Y \ldots$ and we denote their types as $X : \tau$. When the type is not
important, we omit type declarations. We assume that symbols $\itrue,
\ifalse,\inot,\iand, \ior,\iimplies,\iequiv$ with their standard meanings and type declarations are elements of
$\Sigma$. Furthermore, we assume that polymorphic symbols $\iforall$ and $\iexists$
with type declarations $\forallty{\alpha} (\alpha \rightarrow o) \rightarrow o$
and $\ieq \; : \forallty{\alpha} {\alpha \rightarrow \alpha \rightarrow o}$ are
in $\Sigma$, with their standard meanings. All these symbols are called \emph{logical
symbols}. We use infix notation for binary logical symbols.

\looseness=-1 Terms are defined inductively as follows. Free ($X : \tau$)  and bound ($x : \tau$) variables   are
terms of type $\tau$. If $\cst{f} : \forallty{\tuplen{\alpha}} \tau$ is in
$\Sigma$ and $\tuplen{\upsilon}$ is a tuple of types, called type arguments, then
$\cst{f}\typeargs{\tuplen{\upsilon}}$ (written as $\cst{f}$ if $n=0$, or if type
arguments can be inferred from the context) is a term of type $\tau \{
\tuplen{\alpha} \mapsto \tuplen{\upsilon} \}$, called constant. If $x$ is a bound variable
of type $\tau$ and $s$ is a term of type $\upsilon$ then $\lamx{s}$ is a term of type
$\tau \rightarrow \upsilon$. If $s$ and $t$ are of type $\tau \rightarrow \upsilon$ and
$\tau$, respectively, then $s \, t$ is a term of type $\upsilon$. We call terms of
Boolean type ($o$) \emph{formulas} and denote them by $f,g,h, \ldots$; we use
$p,q,r, \ldots$ for free variables whose result type is $o$ and
$\cst{p},\cst{q},\cst{r}$ for constants with the same result type.
%
Unless stated otherwise, we view terms as
$\alpha\beta\eta$-equivalence classes, with $\eta$-long $\beta$-reduced form as
the representative.

\looseness=-1
Given a formula $f$ we call its Boolean subterm $f|_p$ a \emph{top-level
Boolean} if for all proper prefixes $q$ of $p$, the head of $f|_q$ is a logical
constant. Otherwise, we call it a \emph{nested Boolean}. For example, in the formula
$f = \cst{h} \, \cst{a} \ieq \cst{g} \, (\cst{p} \iimplies \cst{q}) \llor
\inot\cst{p}$, $f|_1$ and $f|_2$ are top-level Booleans, while $f|_{1.2.1}$ is
a nested Boolean, as well as its subterms. First-order logic allows only top-level Booleans, whereas nested Booleans are characteristic for
higher-order logic.