### 2022-03-04 ehoh2/ehoh2.tex

\chapter{Extending a Brainiac Prover to Higher-Order Logic}
\setheader{Extending a Brainiac Prover to Higher-Order Logic}
\label{ch:ehoh2}