### 2022-03-05 bools/bools.tex
more

\chapter{Boolean Reasoning in a Higher-Order Superposition Prover}
\setheader{Boolean Reasoning in a Higher-Order Superposition Prover}
\label{ch:bools}

\blfootnote{In this work I implemented and evalauted most of the described approaches.
Visa Nummelin implemented the FOOL preprocessing module.}

\authors{Joint work with Visa Nummelin}

\begin{abstract}
    We present a pragmatic approach to extending a Boolean-free higher-order
    superposition calculus to support Boolean reasoning. Our approach extends
    inference rules that have been used only in a first-order setting, uses some
    well-known rules previously implemented in higher-order provers, as well as new rules.
    We have implemented the approach in the Zipperposition
    theorem prover. The evaluation shows highly competitive performance of our approach
    and clear improvement over previous techniques.
\end{abstract}

\newpage

\section{Introduction} 
\label{sect:bool:introduction}

In the last decades, automatic theorem provers have been used
successfully as backends to ``hammers'' in proof assistants
\cite{ku-15-holyhammer, pb-12-sh} and to software
verifiers \cite{fp-13-why3}. Most advanced provers, such as CVC4
\cite{cbetal-11-cvc4}, E \cite{scv-19-e23}, and Vampire
\cite{lkav-13-vampire}, are based on first-order logic, whereas
most frontends that use them are based on versions of higher-order logic. Thus,
there is a large gap in expressiveness between front- and backends.
%
\looseness=-1
This gap is bridged using well-known translations from higher-order to
first-order logic \cite{ar-70-hol,
mp-08-trans}. However, as shown in Chapter \ref{ch:ehoh}, translations are usually less efficient than native support. The
distinguishing features of higher-order logic used by proof assistants that the
translation must eliminate include
$\lambda$-binders, function extensionality---the property that functions are
equal if they agree on every argument, described by the axiom $\iforall xy. \allowbreak\, (\iforall z. \, x \, z \ieq y \,z)
\iimplies x \ieq y$, and formulas occurring as arguments of function symbols \cite{mp-08-trans}.

\looseness=-1
A group of authors including Vukmirovi\'c\ \cite{bbtvw-21-sup-lam} designed an
extension of superposition for extensional Boolean-free higher-order logic. The extension removes the need to translate the
first two above mentioned features of higher-order logic. Kotelnikov et al.\
\cite{kotelnikov-15-fool,kotelnikov-16-fool} extended the language of
first-order logic to support the third feature of higher-order logic that
requires translation. They described two approaches: One based on calculus-level
treatment of Booleans and the other, which requires no changes to the calculus,
based on preprocessing.

\looseness=-1 To fully bridge the gap between higher-order and first-order tools,
we combine the two approaches: we use the efficient higher-order superposition
calculus and extend it with inference rules that reason with
Boolean terms. In early work, Kotelnikov et al.\ 
\cite{kotelnikov-15-fool} have described a \emph{FOOL paramodulation}
rule that, under some order requirements, removes the need for the axiom
describing the Boolean domain---$\iforall p. \; p \ieq \itrue \llor p
\ieq \ifalse$. In this approach, it is assumed that a problem with formulas occurring as
arguments of symbols is translated to first-order logic.

\looseness=-1
The backbone of our approach is based on an extension of this rule to higher-order
logic: We do not translate away any Boolean structure that is nested inside
non-Boolean terms and allow our rule to hoist the nested Booleans to the
literal level. Then, we clausify the resulting formula (i.e., a clause containing formulas in literals) using a new rule.

\looseness=-1 An important feature that we inherit by building on top of Bentkamp
et al. \cite{bbtvw-21-sup-lam} is support for (function) extensionality. Moving to higher-order
logic with Booleans also means that we need to consider \emph{Boolean extensionality}: $\iforall pq. \, (p
\iequiv q) \iimplies p \ieq q$. We extend the rules of Bentkamp et al.\ 
that treat function extensionality to also treat Boolean extensionality.

\looseness=-1
Rules that extend the two orthogonal approaches form the basis of our support
for Boolean reasoning (Sect.~\ref{sect:bool:native}). In addition, we have
implemented rules that are inspired by the ones used in the
higher-order provers Leo-III \cite{sb-21-leo3} and Satallax
\cite{cb-12-satallax}, such as elimination of Leibniz equality, primitive
instantiation and treatment of choice operator \cite{pa-01-classical-ty-thy}. We
have also designed new rules that use higher-order unification to resolve
Boolean formulas that are hoisted to literal level, delay clausification of non-atomic
literals, reason about formulas under $\lambda$-binders, and many others. Even
though the rules we use are inspired by the ones of refutationally complete
higher-order provers, we do not guarantee completeness of our extension of
$\lambda$-superposition.

\looseness=-1
We compare our native approach with two alternatives which are based on
preprocessing (Sect.~\ref{sect:bool:alternative}). First, we compare it to  an
axiomatization of the theory of Booleans. Second, inspired by work of
Kotelnikov et al.\ \cite{kotelnikov-16-fool}, we implemented the preprocessing approach that does not
require introduction of Boolean axioms.
We discuss some examples, coming from TPTP \cite{gs-17-tptp}, that
illustrate advantages and disadvantages of our approach (Sect.~\ref{sect:bool:examples}).

Our approach is implemented in the Zipperposition theorem prover
\cite{sc-15-simon-phd,sc-supind-17}. This prover was used by Bentkamp et al.\
to implement their higher-order superposition calculus. We further extend
their implementation.

\looseness=-1
We performed an extensive evaluation of our approach (Sect.~\ref{sect:bool:eval}).
In addition to evaluating different configurations of our new rules, we have
compared them to full higher-order provers CVC4, Leo-III,  Satallax and Vampire.
The results suggest that it is beneficial to natively support Boolean reasoning
-- the approach outperforms preprocessing-based approaches. Furthermore, it is
very competitive with state-of-the-art higher order provers. We discuss the differences between our approach and the
approaches we base on, as well as related approaches (Sect.~
\ref{sect:bool:discussion}).

\section{Background} 
\label{sect:bool:background}

We base our work (and parts of the following text) on Bentkamp et al.'s \cite{bbtvw-21-sup-lam} extensional
polymorphic clausal higher-order logic. We extend the syntax of this logic by
adding logical connectives to the signature. The semantic of the logic
is extended by interpreting Boolean type $o$ as a two-element domain. This
amounts to extending Bentkamp et al's fragment of higher-order logic to
full-higher order logic (HOL). Taking a different perspective, this logic is an extension of the one
presented in Chapter \ref{ch:ehoh} with $\lambda$ abstraction, polymorphic
types, and native logical connectives. For reference, we provide precise definition of this polymorphic logic.

\looseness=-1
A signature is a quadruple $(\Sigmaty, \Vty, \Sigma, \VV)$ where $\Sigmaty$ is a
set of type constructors, $\Vty$ is a set of type variables and $\Sigma$ and
$\VV$ are sets of constants and term variables, respectively. We require nullary
type constructor $o$, as well as binary constructor $\rightarrow$
to be in $\Sigmaty$. A type $\tau, \upsilon$ is either a type variable $\alpha \in
\Vty$ or of the form $\kappa(\tau_1, \ldots \tau_n)$ where $\kappa$ is an
$n$-ary type constructor. We write $\kappa$ for $\kappa()$, $\tau \rightarrow
\upsilon$ for $\rightarrow(\tau, \upsilon)$, and we drop parentheses to shorten 
$\tau_1 \rightarrow (\cdots \rightarrow (\tau_{n-1} \rightarrow \tau_n) \cdots)$ into $\tau_1 \rightarrow \cdots \rightarrow
\tau_n$. Each symbol in $\Sigma$ is
assigned a type declaration of the form $\forallty{\tuplen{\alpha}} \tau$ where all variables
occurring in $\tau$ are among $\tuplen{\alpha}$.

\looseness=-1
Function symbols $\cst{a}, \cst{b}, \cst{f}, \cst{g}, \ldots$ are elements of
$\Sigma$; their type declarations are written as $\cst{f} :
\forallty{\tuplen{\alpha}} \tau$. Set $\VV$ contains both free and bound variables, following
the distinction introduced in Sect.~\ref{sec:pre:hol}. Free term variables from the set $\VV$ are written
$F,G,X,Y \ldots$ and we denote their types as $X : \tau$. Bound term variables are written $x,y,z,\ldots$,
and their types are similarly denoted. When the type is not
important, we omit type declarations. We assume that symbols $\itrue,
\ifalse,\inot,\iand, \ior,\iimplies,\iequiv$ with their standard meanings and type declarations are elements of
$\Sigma$. Furthermore, we assume that polymorphic symbols $\iforall$ and $\iexists$
with type declarations $\forallty{\alpha} (\alpha \rightarrow o) \rightarrow o$
and $\ieq \; : \forallty{\alpha} {\alpha \rightarrow \alpha \rightarrow o}$ are
in $\Sigma$, with their standard meanings. All these symbols are called \emph{logical
symbols}. We use infix notation for binary logical symbols.

\looseness=-1 Terms are defined inductively as follows. Free ($X : \tau$)  and bound ($x : \tau$) variables   are
terms of type $\tau$. If $\cst{f} : \forallty{\tuplen{\alpha}} \tau$ is in
$\Sigma$ and $\tuplen{\upsilon}$ is a tuple of types, called type arguments, then
$\cst{f}\typeargs{\tuplen{\upsilon}}$ (written as $\cst{f}$ if $n=0$, or if type
arguments can be inferred from the context) is a term of type $\tau \{
\tuplen{\alpha} \mapsto \tuplen{\upsilon} \}$, called constant. If $x$ is a bound variable
of type $\tau$ and $s$ is a term of type $\upsilon$ then $\lamx{s}$ is a term of type
$\tau \rightarrow \upsilon$. If $s$ and $t$ are of type $\tau \rightarrow \upsilon$ and
$\tau$, respectively, then $s \, t$ is a term of type $\upsilon$. We call terms of
Boolean type ($o$) \emph{formulas} and denote them by $f,g,h, \ldots$; we use
$P,Q,R, \ldots$ for free variables whose result type is $o$ and
$\cst{p},\cst{q},\cst{r}$ for constants with the same result type.
%
Formulas whose top-level symbol is not logical are called \emph{atoms}.
Unless stated otherwise, we view terms as
$\alpha\beta\eta$-equivalence classes, with $\eta$-long $\beta$-reduced form as
the representative.

\looseness=-1
Given a formula $f$ we call its Boolean subterm $f|_p$ a \emph{top-level
Boolean} if for all proper prefixes $q$ of $p$, the head of $f|_q$ is a logical
constant. Otherwise, we call it a \emph{nested Boolean}. For example, in the
formula $f = \cst{h} \, \cst{a} \ieq \cst{g} \, (\cst{p} \iimplies \cst{q})
\llor \inot\cst{p}$, $f|_1$ and $f|_2$ are top-level Booleans, while
$f|_{1.2.1}$ is a nested Boolean, as well as its subterms. First-order logic
allows only top-level Booleans, whereas nested Booleans are characteristic for
higher-order logic.

\section{The Native Approach} 
\label{sect:bool:native}

\looseness=-1
Zipperposition already had some support for Boolean reasoning before we
started extending the calculus of Bentkamp et al. In this section, we first
describe the internals of Zipperposition responsible for reasoning with
Booleans. We continue by describing \NumberOK{15} rules that we have
implemented. For ease of presentation we divide them in three categories.
\subsection{Support for Booleans in Zipperposition}
\label{subsect:bool:zip-bools}

As mentioned in Chapter~\ref{ch:unif}, Zipperposition is an open source prover
written in OCaml. From its inception, it was designed as a prover that supports
easy extension of its base superposition calculus to various theories, including
arithmetic, induction and limited support for higher-order logic
\cite{sc-15-simon-phd}.

In Zipperposition, applications are represented in flattened, spine notation. In
addition, Zipperposition uses associativity of $\iand$ and $\ior$ to flatten out
the nested applications of these symbols. For example, terms $\cst{p} \iand
(\cst{q} \iand \cst{r})$ and $(\cst{p} \iand \cst{q}) \iand \cst{r}$ are
internally stored as $\iand \, \cst{p} \, \cst{q} \, \cst{r}$. 
Zipperposition's support for $\lambda$-terms is used to represent quantified nested
Booleans: Formulas $\iforall x. \, \, f$ and $\iexists x. \, f$ are represented as
$\iforall \, (\lamx{f})$ and $\iexists \, (\lamx{f})$. After clausification of the
input problem, no nested Booleans will be modified or renamed using fresh
predicate symbols.

The version of Zipperposition preceding our modifications distinguished between non\-equational
and equational literals. Following E \cite{scv-19-e23}, we
modified Zipperposition to represent all literals equationally: a nonequational
literal $f$ is stored as $f \eq \itrue$, whereas $\neg f$ is stored as $f
\noteq \itrue$. Equations of the form $f \eq \ifalse $ and $f \noteq \ifalse$ are
transformed into $f \noteq \itrue$ and $f \eq \itrue$, respectively.

\subsection{Core Rules}
\label{subsect:bool:core}

\looseness=-1
Kotelnikov et al.\ \cite{kotelnikov-15-fool}, to the best of our
knowledge, pioneered the approach of extending a first-order superposition prover to support
nested Booleans. They call effects of including the axiom $\iforall
p. \; p \eq \itrue \ior p \ieq \ifalse$ a  ``recipe for disaster''. To combat the
explosive behavior of the axiom, they imposed the following two requirements to
the simplification order $\succ$: $\itrue \succ \ifalse$ and
$\itrue$ and $\ifalse$ are two smallest ground terms with respect to $\succ$. If
these requirements are met, there is no self-paramodulation of the clause
and only paramodulation possible is from literal $p \eq \itrue$ of the mentioned axiom
into a Boolean subterm of another clause. Finally, Kotelnikov et al.\ replace
the axiom with the inference rule \emph{FOOL Paramodulation} (\infname{FP}):
\pagebreak[2]
%
$$ \namedinference{FP}{C[f]}{C[\itrue] \vee f \eq \ifalse} $$
%
\looseness=-1
where $f$ is a nested non-variable Boolean subterm of clause $C$, different from
$\itrue$ and $\ifalse$. In addition, they translate the initial problem containing nested
Booleans  to first-order logic without interpreted Booleans; thus, symbols $\itrue$ and $\ifalse$, and type $o$
correspond to proxy symbols and types introduced during the translation.

We created two rules that are syntactically similar to \infname{FP} but
are adapted for higher-order logic with one key distinction -- we do not perform any translation:
%
\begin{align*}
  & \namedinference{Cases}{C[f]}{C[\ifalse] \vee f \eq \itrue}
  && \namedsimp{CasesSimp}{C[f]}{C[\ifalse] \vee f \eq \itrue \quad C[\itrue] \vee f \noteq \itrue}
\end{align*}
%
\looseness=-1
Clearly, the prover that uses the rules should not include them both at the same time.
In addition, since
literals $f \eq \ifalse$ are represented as negative equations $f \noteq \itrue$, which cannot be used to paramodulate from,
we change the first requirement on the order to $\ifalse \succ
\itrue$.

\newcommand{\eqneq}{\mathrel{\dot{\eq}}}
These two rules hoist Boolean subterms $f$ to the literal level; therefore,
some results of \infname{Cases} and \infname{CasesSimp} will have literals of the form $f \eq \itrue$ (or
$f \noteq \itrue$) where $f$ is not an atom. This introduces the need for the rule
called immediate clausification (\infname{IC}):
%
$$ \namedsimp{EC}{C}{D_1 \; \cdots \; D_m} $$
%
We say that a clause is \emph{standard} if all of its literals are of the form $s \eqneq t$,
where $s$ and $t$ are not Booleans or of the form $f \eqneq \itrue$, where the head of $f$
is not a logical symbol and $\eqneq$ denotes $\eq$ or $\not\eq$. The rule \infname{IC}
is applicable if clause $C = L_1 \llor
\cdots \llor L_n$ is not standard.
The resulting clauses $\tuple{D}{m}$ represent
the result of clausification of the formula $\iforall\,\tuple{x}{}. \; L_1 \ior
\cdots \ior L_n$ where $\tuple{x}{}$ are all free variables of $C$.
Using Boolean extensionality, Zipperposition's clausification
algorithm treats Boolean equality as equivalence (i.e., it replaces
$\mathop{\eq}\typeargs{o}$ with $\iimplies$).

\looseness=-1
An advantage of leaving nested Booleans unmodified is that the prover will be able
to prove some problems containing them without using the prolific rules described
above. For example, given two clauses $\cst{f} \, (\cst{p} \, x
\iimplies \cst{p} \, y) \eq \cst{a}$ and $\cst{f} \, (\cst{p} \,
\cst{a} \iimplies \cst{p} \, \cst{b}) \noteq \cst{a}$, the empty clause can
easily be derived without the above rules. A disadvantage of this approach
is that the proving process will periodically be interrupted by expensive calls
to the clausifier.
\pagebreak[2]

\looseness=-1
If implemented naively, rules \infname{Cases} and \infname{CasesSimp} can result in many redundant clauses. Consider the following
example: let $\cst{p} : o \rightarrow o$, $\cst{a} : o$ and consider a clause
set containing $\cst{p} \, (\cst{p} \, (\cst{p} \, (\cst{p} \, \cst{a}))) \eq
\itrue$. Then, the clause $C = \cst{a} \eq \itrue \llor \cst{p} \, \ifalse \eq \itrue$ can be
derived in eight ways using the rules, depending on which nested Boolean subterm was
chosen for the inference. In general, if a clause has a subterm occurrence of the form $\cst{p}^n \, \cst{a}$,
where both $\cst{p}$ and $\cst{a}$ have result type $o$, the clause $\cst{a} \eq \itrue \llor \cst{p} \, \ifalse \eq \itrue$ can be derived in $2^{n-1}$ ways.
% Boolean subterms occurring in different
% eligible literals cause similar explosive behavior.
To combat these issues we
implemented pragmatic restrictions of the rule: only $f$ which is
the leftmost outermost (or innermost) eligible subterm will be considered. With
this modification $C$ can be derived in only one way. Furthermore,
some intermediate conclusions of the rules will not be derived, pruning the search space.

The clausification algorithm by Nonnengart and Weidenbach \cite{nw-01-small-cnf}
eagerly simplifies the input problem using well-known Boolean equivalences before clausifying it. For example,
the formula $\cst p\land\itrue$ is replaced by $\cst p$. To simplify nested Booleans we implemented the rule

\[
\namedsimp{BoolSimp}{C[\sigmaterm{f}]}{C[\sigmaterm{g}]}
\]
where $f \longrightarrow g \in E$ runs over fixed set of rewrite rules $E$,
and $\sigma$ is any substitution. In the current implementation of Zipperposition, $E$ contains the rules
described by Nonnengart and Weidenbach \cite[Sect.~3]{nw-01-small-cnf}. This set contains
the rules describing how each logical symbol behaves when given arguments $\itrue$ or $\ifalse$: for example, 
it includes $(\itrue \iimplies p)~\longrightarrow~p $ and $(p \iimplies \itrue) \longrightarrow \itrue$. %full list visa report Sect 4.2
Leo-III implements a similar rule, called \textsf{simp} \cite[Sect.~4.2.1.]{as-18-phd}.

Our decision to represent negative atoms as negative equations was motivated by
the need to alter Zipperposition's earlier behavior as little as possible. 
Namely, negative atoms were not used as literals that can be used
to paramodulate from, and as such added to the laziness of the superposition calculus.
However, it might be useful to consider unit clauses of the form $f \noteq \itrue$
as $f \eq \ifalse$ to strengthen rewriting. To that end, we have introduced the following
rule:
%
$$ \namedsimp{BoolDemod}{f \noteq \itrue \qquad C[\sigmaterm{f}]}{f \noteq \itrue \qquad C[\ifalse]} $$

\subsection{Higher-Order Considerations}
\label{subsect:bool:core}
\looseness=-1
To achieve refutational completeness of higher-order resolution and similar
calculi it is necessary to instantiate variables with result type $o$,
\emph{predicate variables}, with arbitrary formulas
\cite{as-18-phd,pa-01-classical-ty-thy}. Fortunately, we can approximate the
formulas using a complete set of logical symbols (e.g., $\inot$, $\iforall$, and
$\iand$). Since such an approximation is not only necessary for completeness of
some calculi, but very useful in practice, we implemented the \emph{primitive
instantiation} (PI) rule:
%
$$ \namedinference{PI} {C \llor \lam{\tuple{x}{}}{P \, \tuplen{s}} \, \dot{\eq}
\, t} {\substcl{\{ P
\mapsto f \}}{C \llor \lam{\tuple{x}{}}{P \, \tuplen{s}} \, \dot{\eq} \, t}  } $$
%
where $P$ is a free variable of
the type $\tau_1 \rightarrow \cdots \rightarrow \tau_n \rightarrow o$. 
Choosing a different $f$ that instantiates $P$, we can balance between
explosiveness of approximating a complete set of logical symbols and
incompleteness of pragmatic approaches. We borrow the notion of imitation from
higher-order unification jargon (Sect.~\ref{sec:unif:the-unification-procedure}) %\cite{OUR UNIFICATION PAPER}:
, and we say
that the term $\lam{\tuple{x}{m}}{\cst{f} \, (Y_1 \, \tuple{x}{m}) \cdots (Y_n
\, \tuple{x}{m}) }$ is an \emph{imitation} of constant $\cst{f} : \tau_1
\rightarrow \cdots \rightarrow \tau_n \rightarrow \tau$ for some variable $z$ of type $\nu_1
\rightarrow \cdots \rightarrow \nu_m \rightarrow \tau$. Variables $\tuplen{Y}$
are fresh free variables, where each $Y_i$ has the type $\nu_1 \rightarrow
\cdots \rightarrow \nu_m \rightarrow \tau_i$; variable $x_i$ is of type $\nu_i$.
\pagebreak[2]

\looseness=-1
Rule \infname{PI} was already implemented by Simon Cruanes in Zipperposition,
before we started our modifications. The rule has different modes that generate
sets of possible terms $f$ for $p: \tau_1 \rightarrow \cdots \rightarrow \tau_n
\rightarrow o$: \emph{Full}, \emph{Pragmatic}, and $\mathit{Imit}_{\star}$ where
$\star$ is an element of a set of logical constants $P = \{ \iand, \ior,
\mathop{\ieq}\typeargs{\alpha}, \inot, \iforall, \iexists \}$. Mode \emph{Full}
contains imitations (for $p$) of all elements of $P$. Mode $\emph{Pragmatic}$
contains imitations of $\inot$, $\itrue$ and $\ifalse$; if there exist indices $i,j$
such that $i\ineq j$ and  $\tau_i = \tau_j$, it contains $\lam{\tuplen{x}}{x_i
\ieq x_j}$; if there exist indices $i,j$ such that $i \ineq j$, and $\tau_i =
\tau_j = o$, then it contains $\lam{\tuplen{x}}{x_i \iand x_j}$ and
$\lam{\tuplen{x}}{x_i \ior x_j}$; if for some $i$, $\tau_i = o$, then it
contains $\lam{\tuplen{x}}{x_i}$. Mode $\mathit{Imit}_\star$ contains
imitations of $\itrue$, $\ifalse$ and $\star$ (except for $\mathit{Imit_{\iforall\iexists}}$ which contains imitations
of both $\iforall$ and $\iexists$).

\looseness=-1
While experimenting with our implementation we have noticed some proof patterns
that led us to come up with the following modifications. First, it often
suffices to perform \infname{PI} only on initial clauses -- which is why we
allow the rule to be applied only to the clauses created using at most $k$
generating inferences. Second, if the rule was used in the proof, its premise
is usually only used as part of that inference -- which is why we implemented
a version of \infname{PI} that removes the clause after all possible \infname{PI}
inferences have been performed. We observed that the mode \emph{Imit}$_\star$ is useful in practice since often only a single
approximation of a logical symbol is necessary.

\looseness=-1
Efficiently treating axiom of choice is notoriously difficult for higher-order provers. Andrews formulates
this axiom as $\iforall p.\, (\iexists
x.\, p \, x) \iimplies p \, (\varepsilon \, p)$, where $\varepsilon :
\forallty{\alpha}(\alpha \rightarrow o) \rightarrow \alpha$ denotes the \emph{choice
operator} \cite{pa-01-classical-ty-thy}. After clausification, this axiom becomes $P \, X \noteq \itrue \llor P
\, (\varepsilon \, P) \eq \itrue$. Since term $P \, X$ matches any Boolean
term in the proof state, this axiom is very explosive. Therefore, Leo-III
\cite{sb-21-leo3} deals with the choice operator on the calculus
level. Namely, whenever a clause $C = P \, X \noteq \itrue \llor P \, (\cst{f} \,
P) \eq \itrue$ is chosen for processing, $C$ is removed from the proof state and $\cst{f}$ is
added to set of choice functions $\mathit{CF}$ (which initially contains just
$\varepsilon$). Later, elements of $\mathit{CF}$ will be used to heuristically
instantiate the axiom of choice. We reused the method of recognizing choice
functions, but generalized the rule for creating the instance of
the axiom (assuming $\xi \in \mathit{CF}$):
%
$$\namedinference{Choice}
               {C[\xi \, t]}
               { X \, (t \, Y) \noteq \itrue \llor X \, (t \, (\xi \, (\lam{z}{X \, (t \, z)})) ) \eq \itrue }$$
%

Let $D$ be the conclusion of \infname{Choice}. The fresh variable $X$ in $D$ acts as
arbitrary context around $t$, the chosen instantiation for $P$ from axiom of choice;
the variable $X$ can later be replaced by imitation of logical symbols to create more
complex instantiations of the choice axiom. To generate useful instances early, we create two instances: $\substcl{\{X
\mapsto \lam{z}{z}\}}{D}$ and $\substcl{\{X \mapsto \lam{z}{\inot\, z}\}}{D}$. Then, based on
Zipperposition parameters, $D$ will either be deleted or kept. Note that $D$
will not subsume its instances, since the matching algorithm of Zipperposition is
too weak for this.
\looseness=-1

Most provers natively support extensionality reasoning: Bhayat et al. \cite{br-20-full-sup-w-combs} modify
first-order unification to return unification constraints consisting of pairs of
terms of functional type, whereas Steen relies on the unification rules of
Leo-III's calculus \cite[Section 4.3.3.]{as-18-phd} to deal with extensionality.
Bentkamp et al \cite{bbtvw-21-sup-lam} altered core generating inference rules of the superposition calculus
to support extensionality. Instead of requiring that terms
involved in the inference are unifiable, it is required that they can be
decomposed into \emph{disagreement pairs} such that at least one of the
disagreement pairs is of functional type. Disagreement pairs of terms $s$ and
$t$ of the same type are defined inductively using function $\textsf{dp}$:
$\textsf{dp}(s,t) = \emptyset$ if $s$ and $t$ are equal; $\textsf{dp}(a \,
\tuplen{s}, b \, \tuple{t}{m}) = \{(a \,
\tuplen{s}, b \, \tuple{t}{m})\}$ if $a$ and $b$ are different heads;
$\textsf{dp}(\lam{x}{s}, \lam{y}{t}) = \{ (\lam{x}{s}, \lam{y}{t}) \}$;
$\textsf{dp}(a \, \tuplen{s}, a \, \tuplen{t}) = \bigcup_{i=1}^{n}
\textsf{dp}(s_i, t_i)$. Then the extensionality rules are stated as follows:
$$
\namedinference{AbsSup}
  {s \eq t \llor C \qquad u[s'] \mathrel{\dot{\eq}} v \llor D}
  {\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor  u[t] \mathrel{\dot{\eq}} v \llor C \llor D}} $$
\pagebreak[2]
\begin{align*}
  &\namedinference{AbsER}{s \noteq s' \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \cdots \llor  C}} \\
  &\namedinference{AbsEF}{s \eq t \llor  s' \eq u \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor t \noteq u \llor  s' \eq u \llor C}}
\end{align*}

\noindent\looseness=-1
Rules \infname{AbsSup}, \infname{AbsER}, and \infname{AbsEF} are extensional
versions of superposition, equality resolution and equality factoring
(Sect.~\ref{sec:pre:rules}). By \infname{Abs} we denote the union of these three
rules. In each of the rules, $\sigma$ is a most general unifier of the types of
$s$ and $s'$, and $\textsf{dp}(s\sigma ,s' \sigma) = \{ (s_1, s'_1), \ldots,
(s_n, s'_n)\}$. All side conditions for extensional rules are the same as for
the standard rules, except that condition that $s$ and $s'$ are unifiable is
replaced by the condition that at least one $s_i$ is of functional type and that
$n>0$. This rule is easily extended to support Boolean extensionality by
requiring that at least one $s_i$ is of functional or type $o$, and adding the
condition ``$\textsf{dp}(f, g) = \{(f,g)\}$ if $f$ and $g$ are different
formulas'' to the definition of $\textsf{dp}$.

Consider the clause $\cst{f} \, (\inot\cst{p} \ior \inot\cst{q}) \noteq \cst{f} \,
(\inot (\cst{p} \iand \cst{q}))$. This problem is obviously unsatisfiable, since
arguments of $\cst{f}$ on different sides of the disequation are extensionally
equal; however, without \infname{Abs} rules Zipperposition will rely on
\infname{Cases}(\infname{Simp}) and \infname{EC} rules to derive the empty
clause. Rule \infname{AbsER} will generate~$C = \inot\cst{p} \ior \inot\cst{q} \noteq
\inot (\cst{p} \iand \cst{q})$. Then, $C$ will get clausified using
\infname{EC}, effectively reducing the problem to $\inot (\inot\cst{p} \ior
\inot\cst{q} \iequiv \inot (\cst{p} \iand \cst{q}))$, which is first-order.

\looseness=-1
Zipperposition restricts \infname{AbsSup} by
requiring that $s$ and $s'$ are not of function or Boolean types. If the terms are of function type, our experience is
that better treatment of function extensionality is to apply fresh free
variables (or Skolem terms, depending on the sign
\cite{bbtvw-21-sup-lam}) to both sides of a (dis)equation to reduce it to
a first-order literal; Boolean extensionality is usually better supported by
applying \infname{EC} on the top-level Boolean term. Thus, for the following
discussion we can assume $s$ and $s'$ are not $\lambda$-abstractions or formulas. Then, $\infname{AbsSup}$ is applicable if $s$
and $s'$ have the same head, and a functional or Boolean subterm. To speed up
retrieval of such terms, we added an index that maps symbols to positions in
clauses where they appear as a head of a term that has a functional or Boolean
subterm. This index will be empty for first-order problems, incurring no
overhead if extensionality reasoning is not needed. One more restriction we implemented is that we do not apply \infname{Abs} rules if all
disagreement pairs have at least one side whose head is a variable; those will
be dealt with more efficiently using standard, non-extensional, versions of the
rules. We also eagerly resolve literals $s_i \not\eq s'_i$ using at most one
unifier returned by terminating, pragmatic variant of unification algorithm described in 
Sect.~\ref{sec:unif:the-unification-procedure}.

Expressiveness of higher-order logic allows users to define equality using a single axiom,
called Leibniz equality \cite{pa-01-classical-ty-thy}:
%
$ \iforall xy. \,
   (\iforall p. \, p \, x \iimplies p \, y) \iimplies x \ieq y$.
%
Leibniz equality often appears in TPTP problems. Since modern provers have the native support
for equality, it is usually beneficial to recognize and replace occurrences of Leibniz equality.

Before we began our modifications, Zipperposition had a powerful rule that
recognizes clauses that contain variations of Leibniz equality and instantiates
them with native equality. This rule was designed by Simon Cruanes, and to the
best of our knowledge, it has not been documented so far. With his permission
we describe this rule as follows:
\pagebreak[2]
%
$$\namedinference{ElimPredVar}{P \, \overline{s}_n^1 \eq \itrue \llor \cdots \llor
    P \, \overline{s}_n^i \eq \itrue \llor P \, \overline{t}_n^1 \noteq \itrue \llor
    \cdots \llor P \, \overline{t}_n^j \noteq \itrue \llor C} 
    {\sigmacl{P \, \overline{s}_n^1
    \eq \itrue \llor \cdots \llor P \, \overline{s}_n^i \eq \itrue  \llor C}} $$
%
where $P$ is a free variable, $P$ does not occur in any $s_k^l$ or $t_k^l$, or
in $C$; $\sigma$ is defined as $\{ P \mapsto \lam{\tuplen{x}}{\pmb{\bigvee}_{k=1}^{j}
(\pmb{\bigwedge}_{l=1}^{n} x_l \eq t^k_l ) } \}$.

\looseness=-1
To better understand how this rule removes variable-headed negative literals,
consider the clause $C = P \, \cst{a}_1 \, \cst{a}_2 \eq \itrue \llor P \,
\cst{b}_1 \, \cst{b}_2 \not\eq \itrue \llor P \, \cst{c}_1 \, \cst{c}_2 \not\eq
\itrue$. Since all side conditions are fulfilled, the rule \infname{ElimPredVar}
will generate $\sigma = \{ P \mapsto \lam{xy}{(x \ieq \cst{b}_1
\iand y \eq \cst{b}_2) \ior (x \ieq \cst{c}_1 \iand y \ieq \cst{c}_2)  } \}$.
After applying $\sigma$ to $C$ and subsequent $\beta$-reduction, negative literal
$  P \, \cst{b}_1 \, \cst{b}_2 \not\eq \itrue$ will reduce to 
$ (\cst{b}_1 \ieq \cst{b}_1 \iand \cst{b}_2 \ieq \cst{b}_2) \ior (\cst{b}_1 \ieq \cst{c}_1 \iand \cst{b}_2 \eq \cst{c}_2) \noteq \itrue $,
which is equivalent to $\ifalse$. Thus, we can remove this literal and all negative literals
of the form $P \, \overline{t}_n \noteq \itrue$ from $C$ and apply $\sigma$ to the remaining ones.

The previous rule removes all variables occurring in disequations in one
attempt. We implemented two rules that behave more lazily, inspired by the ones present in Leo-III and
Satallax:
%
\begin{align*}
&\namedinference{ElimLeibniz$+$}
{P \, \tuplen{s} \eq \itrue \llor P \, \tuplen{t} \noteq \itrue \llor C}
{\sigmacl{s_i \eq t_i \llor C}}
&&
\namedinference{ElimLeibniz$-$}
{P \, \tuplen{s} \noteq \itrue \llor P \, \tuplen{t} \eq \itrue \llor C}
{\substcl{\sigma'}{s_i \eq t_i \llor C}}&
\end{align*}
%
\looseness=-1
where $P$ is a free variable, $P$ does not occur in $t_i$, $\sigma =
\{ P \mapsto \lam{\tuplen{x}}{x_i \eq t_i} \}$ and
$\sigma' = \{ P \mapsto \lam{\tuplen{x}}{\neg (x_i \eq t_i)} \}$. This
rule differs from \infname{ElimPredVar} in three ways. First, it acts on
occurrences of variables in both positive and negative literals. Second,
due to its simplicity, it usually does not require \infname{IC} as the following step.
Third, it imposes much weaker conditions on $P$. However, removing all negative
variables in one step might improve performance.  Coming back to example of the clause $C = P \, \cst{a}_1 \, \cst{a}_2 \eq \itrue \llor P \,
\cst{b}_1 \, \cst{b}_2 \not\eq \itrue \llor P \, \cst{c}_1 \, \cst{c}_2 \not\eq
\itrue$, we can apply \infname{ElimLeibniz$+$} using the substitution $\sigma = \{ P \mapsto \lam{xy}{x \eq \cst{b_1}} \}$
to obtain the clause $C' = \cst{a}_1 \eq \cst{b}_1 \llor \cst{a}_1 \not\eq \cst{c}_1$.

% \subsection{Additional Rules}
% \label{subsect:bool:core}

% Zipperposition's unification algorithm 
% %\cite{OUR UNIFICATION PAPER}
% uses flattened representation of terms with logical operators $\land$ and $\lor$
% for heads to unify terms that are not unifiable modulo $\alpha\beta\eta$-equivalence, but
% are unifiable modulo associativity and commutativity of $\land$ and $\lor$. Let
% $\diamond$ denote either $\land$ or $\lor$. When the unification algorithm is given
% two terms $\diamond \, \tuplen{s}$ and $\diamond \, \tuplen{t}$, where neither
% of $\tuplen{s}$ nor $\tuplen{t}$ contain duplicates, it performs the
% following steps: First, it removes all terms that appear in both
% $\tuplen{s}$ and $\tuplen{t}$ from the two argument tuples.
% %Let $\tuple{s}{m}$ and $\tuple{t}{m}$ denote the results of previous step.
% Next, the remaining terms are sorted first by their head term and then their weight. Finally,
% an attempt is made to unify sorted lists pairwise.
% %
% As an example, consider the problem of unifying the pair $\big({\land} \, (\cst{p} \,
% \cst{a}) \; (\cst{q} \, (\cst{f} \, \cst{a})), \; {\land} \; (\cst{q} \, (\cst{f} \,
% \cst{a})) \; (r \, (\cst{f} \, (\cst{f} \, \cst{a})))\bigr)$ where $r$ is a free
% variable. If the arguments of $\land$ are simply sorted as described above, we would
% try to unify $\cst{p} \, \cst{a}$ with $\cst{q} \, (\cst{f} \, \cst{a})$, and
% fail to find a unifier. However, by removing term  $\cst{q} \, (\cst{f} \,
% \cst{a})$ from the argument lists, we will be left with the problem
% $(\cst{p} \, \cst{a},  r \, (\cst{f} \, (\cst{f} \, \cst{a})))$ which has a unifier.