### 2022-03-28 title/title.tex
concl

% For VSCode
% !TEX root = ../dissertation.tex
\begin{titlepage}

% \begin{center}

% %% Extra whitespace at the top.
% \vspace*{2\bigskipamount}

% %% Print the title.
% {\makeatletter
% \titlestyle\bfseries\LARGE\@title
% \makeatother}

% %% Print the optional subtitle.
% {\makeatletter
% \ifx\@subtitle\undefined\else
%     \bigskip
%     \titlefont\titleshape\Large\@subtitle
% \fi
% \makeatother}

% \end{center}

\cleardoublepage
\thispagestyle{empty}

\begin{center}

%% The following lines repeat the previous page exactly.

\vspace*{2\bigskipamount}

%% Print the title.
{\makeatletter
\titlestyle\bfseries\LARGE\@title
\makeatother}

%% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
    \bigskip
    \titlefont\titleshape\Large\@subtitle
\fi
\makeatother}

%% Uncomment the following lines to insert a vertically centered picture into
%% the title page.
%\vfill
%\includegraphics{title}
\vfill

%% Apart from the names and dates, the following text is dictated by the
%% promotieregelement.

{\Large\titlefont\bfseries Proefschrift}

\bigskip
\bigskip

ter verkrijging van de graad van doctor

aan Vrije Universiteit Amsterdam,

op gezag van de Rector Magnificus prof.dr.~J.J.G.~Geurts,

voorzitter van het College voor Promoties,

in het openbaar te verdedigen

op [onbekend]

\bigskip
\bigskip

door

\bigskip
\bigskip

%% Print the full name of the author.
\makeatletter
{\Large\titlefont\bfseries\@firstname\ \titleshape{\MakeUppercase{\@lastname}}}
\makeatother

\bigskip
\bigskip

Master of Science in Computer Science, \\
Vrije Universiteit Amsterdam, Nederland,

geboren te Belgrado, Servi\"e.

%% Extra whitespace at the bottom.
\vspace*{2\bigskipamount}

\end{center}

\clearpage
\thispagestyle{empty}

%% The following line is dictated by the promotieregelement.
\noindent Dit proefschrift is goedgekeurd door de

%% List the promotors (supervisors).
\medskip\noindent
\begin{tabular}{ll}
    promotor: & prof.dr. W. J. Fokkink \\
    copromotoren: & dr. J. C. Blanchette \\
                  & prof.$\,$dr. S. Schulz
\end{tabular}

\bigskip
\noindent Samenstelling promotiecommissie:

%% List the committee members, starting with the Rector Magnificus and the
%% promotor(s) and ending with the reserve members.
\medskip\noindent
\begin{tabular}{p{4.5cm}l}

prof.$\,$dr. Mark Hoogendoorn & Vrije Universiteit Amsterdam \\
dr. Konstantin Korovin & The University of Manchester \\
dr. Ruzica Piskac & Yale University \\
dr. Alexander Steen & University of Greifswald \\
dr. Martin Suda & Czech Technical University in Prague \\
prof.$\,$dr. Geoff Sutcliffe & University of Miami \\

\end{tabular}

%% Include the following disclaimer for committee members who have contributed
%% to this dissertation. Its formulation is again dictated by the
%% promotieregelement.
%\medskip
%\noindent  %Prof.\ Dr.\ D.\ Spinellis has contributed to the creation of this thesis.

\medskip
\medskip
% TODO Include http://www.win.tue.nl/ipa/?page_id=309
\noindent The work in the thesis has been carried out under the auspices of the research school IPA
(Institute for Programming research and Algorithmics).

\medskip
%% Here you can include the logos of any institute that contributed financially
%% to this dissertation.
\vfill
\begin{center}
    \includegraphics[height=0.5in]{title/logos/vulogo}
    \hspace{2em}
    %\includegraphics[height=0.5in]{title/logos/casimir} \\
    \includegraphics[height=0.5in]{title/logos/erclogo}
    \\ \vspace{0.5cm}
    \includegraphics[height=0.5in]{title/logos/ipa}
\end{center}
\vfill

\noindent
\begin{tabular}{@{}p{0.2\textwidth}@{}p{0.8\textwidth}}
  \textit{Keywords:} &  \\[\medskipamount]
      \textit{Printed by:} &  \\[\medskipamount]
      \textit{Cover:} &  \\[\medskipamount]
      \textit{Style:} & TU Delft House Style, with modifications by Moritz Beller \\& \url{https://github.com/Inventitech/phd-thesis-template} \\[\medskipamount]
\end{tabular}

\medskip
\medskip
\noindent The author set this thesis in \LaTeX\xspace using the Libertinus Serif, Inter and Inconsolata fonts.

\vspace{\bigskipamount}

% Copyrighting this is stupid, questionable, and probably illegal, because large parts of the
% thesis have already been published with the copyright resigning with the publisher.
%\noindent Copyright \textcopyright\ 2015 by A.~Einstein

%% Uncomment the following lines if this dissertation is part of the Casimir PhD
%% Series, or a similar research school.
%\medskip
%\noindent Casimir PhD Series, Delft-Leiden 2015-01

%\medskip
\noindent ISBN \ldots

\medskip
\noindent An electronic version of this dissertation is available at \\
\url{http://repository.tudelft.nl/}.

\end{titlepage}