### 2022-08-25 title/title.tex
more

% For VSCode
% !TEX root = ../dissertation.tex
\begin{titlepage}

% \begin{center}

% %% Extra whitespace at the top.
% \vspace*{2\bigskipamount}

% %% Print the title.
% {\makeatletter
% \titlestyle\bfseries\LARGE\@title
% \makeatother}

% %% Print the optional subtitle.
% {\makeatletter
% \ifx\@subtitle\undefined\else
%     \bigskip
%     \titlefont\titleshape\Large\@subtitle
% \fi
% \makeatother}

% \end{center}

\cleardoublepage
\thispagestyle{empty}

\begin{center}

%% The following lines repeat the previous page exactly.

VRIJE UNIVERSITEIT

\vspace*{4\bigskipamount}

%% Print the title.
{\makeatletter
\titlestyle\bfseries\LARGE\@title
\makeatother}

%% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
    \bigskip
    \titlefont\titleshape\Large\@subtitle
\fi
\makeatother}

%% Uncomment the following lines to insert a vertically centered picture into
%% the title page.
%\vfill
%\includegraphics{title}
% \vfill
\vspace*{5\bigskipamount}

%% Apart from the names and dates, the following text is dictated by the
%% promotieregelement.

{ACADEMISCH PROEFSCHRIFT}

\bigskip
\bigskip

ter verkrijging van de graad Doctor aan

de Vrije Universiteit Amsterdam,

op gezag van de rector magnificus

prof.dr. J.J.G. Geurts,

in het openbaar te verdedigen

ten overstaan van de promotiecommissie

van de Faculteit der Bètawetenschappen

op dinsdag 18 oktober 2022 om 13.45 uur

in een bijeenkomst van de universiteit,

De Boelelaan 1105

\bigskip
\bigskip

door

\smallskip

Petar Vukmirovi\'c

\smallskip

geboren te Belgrado, Servi\"e

%% Extra whitespace at the bottom.
\vspace*{2\bigskipamount}

\end{center}

\clearpage
\thispagestyle{empty}

%% The following line is dictated by the promotieregelement.

%% List the promotors (supervisors).
\medskip\noindent
\begin{tabular}{ll}
    promotor: & prof.$\,$dr. W. J. Fokkink \\[4\jot]
    copromotoren: & dr. J. C. Blanchette \\
                  & prof.$\,$dr. S. Schulz \\[4\jot]
    promotiecommissie: & prof.$\,$dr. M. Hoogendoorn  \\
                      & dr. K. Korovin               \\
                      & dr. R. Piskac                \\
                      & dr. A. Steen                 \\
                      & dr. M. Suda                  \\
                      & prof.$\,$dr. G. Sutcliffe    \\

\end{tabular}

% \bigskip
% \noindent Promotiecommissie:

% %% List the committee members, starting with the Rector Magnificus and the
% %% promotor(s) and ending with the reserve members.
% \medskip\noindent
% \begin{tabular}{p{4.5cm}l}

% prof.$\,$dr. M. Hoogendoorn & Vrije Universiteit Amsterdam \\
% dr. K. Korovin & The University of Manchester \\
% dr. R. Piskac & Yale University \\
% dr. A. Steen & University of Greifswald \\
% dr. M. Suda & Czech Technical University in Prague \\
% prof.$\,$dr. G. Sutcliffe & University of Miami \\

% \end{tabular}

%% Include the following disclaimer for committee members who have contributed
%% to this dissertation. Its formulation is again dictated by the
%% promotieregelement.
%\medskip
%\noindent  %Prof.\ Dr.\ D.\ Spinellis has contributed to the creation of this thesis.

\newpage
% TODO Include http://www.win.tue.nl/ipa/?page_id=309
\noindent The work in the thesis has been carried out under the auspices of the research school IPA
(Institute for Programming research and Algorithmics).

\medskip
%% Here you can include the logos of any institute that contributed financially
%% to this dissertation.
\vfill
\begin{center}
    \includegraphics[height=0.5in]{title/logos/vulogo}
    \hspace{2em}
    %\includegraphics[height=0.5in]{title/logos/casimir} \\
    \includegraphics[height=0.5in]{title/logos/erclogo}
    \\ \vspace{0.5cm}
    \includegraphics[height=0.5in]{title/logos/ipa}
\end{center}
\vfill

\noindent
\begin{tabular}{@{}p{0.2\textwidth}@{}p{0.8\textwidth}}
  \textit{Keywords:} & automatic theorem proving, higher-order logic, unification  \\[\medskipamount]
      \textit{Printed by:} &  \\[\medskipamount]
      \textit{Cover:} &  Marina Abramović, Rhythm 10 (The Biography), Cibachrome Print,
                         1997, © Marina Abramović,
                         Courtesy of the Marina Abramović Archives \\[\medskipamount]
      \textit{Style:} & TU Delft House Style, with modifications by Moritz Beller \\& \url{https://github.com/Inventitech/phd-thesis-template} \\[\medskipamount]
\end{tabular}

\medskip
\medskip
\noindent The author set this thesis in \LaTeX\xspace using the Libertinus Serif, Inter and Inconsolata fonts.

\vspace{\bigskipamount}

% Copyrighting this is stupid, questionable, and probably illegal, because large parts of the
% thesis have already been published with the copyright resigning with the publisher.
%\noindent Copyright \textcopyright\ 2015 by A.~Einstein

%% Uncomment the following lines if this dissertation is part of the Casimir PhD
%% Series, or a similar research school.
%\medskip
%\noindent Casimir PhD Series, Delft-Leiden 2015-01

%\medskip
% \noindent ISBN \ldots

\medskip
% \noindent An electronic version of this dissertation is available at \\
% \url{http://repository.tudelft.nl/}.

\end{titlepage}